---
title: "Wilcoxons - Week 3, Lecture 1"
author: "Hector Gavilanes"
format: html
self-contained: true
editor: source
---

```{r setup, echo = TRUE, warnings = FALSE, message = FALSE}

library(tidyverse)
library(gsheet)
library(formattable)
library(gridExtra)
```

## Question 1:

**Recall the first dataset from last week. A large credit union awards quarterly bonuses to its employees. To ensure fairness across departments, data on the bonuses are regularly reviewed for equality. Consider the data in the following Google Sheet, which includes the bonus percentage for each employee in two departments.**

```{r, echo = TRUE}
# clear environment
rm(list = ls())

# extract data
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1q0Ouhr6hGBDp-5oyQuYGcsWTLrsayywVHJS7o17Sm6Y/edit?usp=sharing"))

# transform data to Bonus dataframe
bonus_data <- data %>% 
  # delete empty rows
  drop_na()

# ---CUSTOM COLORS---
custom_green_dark = "#DeF7E9"
custom_green_light = "#71CA97"
custom_gray = "#3D3D3D"
custom_orange = "#FC7307"
custom_purple = "#5A18C9"
custom_tan_light = "#F7E3D4"
custom_density_hue = c(custom_purple,custom_orange)
# ---FUNCTIONS---
# Change default table format
change_table_format <- function(data_table){
  # local variables
  first_column <- data_table[1]
  dataframe_length <- NCOL(data_table)
  # format
  formattable(data_table, align = c("l",rep("c", dataframe_length - 1)),
              list(first_column = formatter("span", style = ~ style(color = custom_gray, font.weight = "bold")),
              area(row = 1, col = 2:dataframe_length) ~ color_tile(custom_green_dark, custom_green_light)
              )#/end list
              )#/end formattable
}
```

#### a. (From last week!) Use the appropriate *t*-test to determine if the two departments are given different bonus amounts. Test at the &alpha;=0.01 level.

```{r, echo=TRUE}
# two-sided, two-sample t-test
t.test(bonus_data$Bonus ~ bonus_data$Department,
       var.equal = TRUE,
       )
```

### Hypothesis
H0: $\mu Credit = \mu Mortgage$\
H1: $\mu Credit \ne \mu Mortgage$\

### Test
t0 = 4.04

### p-Value
p < 0.001

## Conclusion
Reject H0 at the $\alpha$=0.01 level. There is sufficient evidence to suggest that the two departments are given different bonuses.

#### b. Use the appropriate nonparametric test to determine if the two departments are given different bonus amounts. Test at the &alpha;=0.01 level.

```{r, echo=TRUE}
# nonparametric test for independent groups
wilcox.test(bonus_data$Bonus ~ bonus_data$Department)
```

### Hypothesis
H0: $\mu Credit = \mu Mortgage$\
H1: $\mu Credit \ne \mu Mortgage$\

### Test
$\T_0$ = 683

### p-Value
p < 0.001

## Conclusion
Reject H0 at the $\alpha$=0.01 level. There is sufficient evidence to suggest that the two departments are given different bonuses.

#### c. Assess the assumption(s) associated with the *t*-test in part a. In your assessment, determine if the *t*-test or the corresponding nonparametric test should be reported.

### Assumptions
```{r}
credit_qq <- bonus_data %>% 
  filter(Department == "Credit") %>% 
  ggplot(aes(sample = Bonus)) +
    stat_qq(size=3) +
    stat_qq_line() +
    theme_minimal() +
    xlab("Theoretical") +
    ylab("Credit Sample")
credit_qq

mortgage_qq <- bonus_data %>% 
  filter(Department == "Mortgage") %>% 
  ggplot(aes(sample = Bonus)) +
    stat_qq(size=3) +
    stat_qq_line() +
    theme_minimal() +
    xlab("Theoretical") +
    ylab("Mortgage Sample")
mortgage_qq
```
### Frequency Table
```{r}
# mean by department
groups_mean <- bonus_data %>% 
  group_by(Department) %>% 
  summarise(Mean = mean(Bonus),
            Median = median(Bonus))
change_table_format(groups_mean)

# bonus frequency
bonus_data %>% 
  group_by(Department) %>%
  ggplot(aes(x = Bonus, colour = Department, fill = Department)) +
  geom_freqpoly(bins = 15)+
  theme_bw()
# bonus density
bonus_data %>% 
  group_by(Department) %>% 
  ggplot(aes(x = Bonus, colour = Bonus, fill = Department)) +
  geom_density(alpha=0.2)+
  geom_vline(data = groups_mean, aes(xintercept=Median, color=Department), linetype="dashed", size = 1.1)+
  xlim(5,13)+
  labs(title = "Credit & Mortgage", subtitle = "Median Bonus")+
  theme(legend.position="bottom")+
  theme_bw()+
  scale_fill_manual(values = custom_density_hue)+
  scale_color_manual(values = custom_density_hue)

```

#### d. Write a brief paragraph explaining your findings to management, based on the test chosen in part c.

## Conclusion
The T-test demonstrated that there was a different of bonuses earned for both groups. Especially towards the right tail, we can find some outliers.

## Question 2: 

**Recall the second dataset from last week. A new 2 hour reading comprehension program has been developed by a local tutoring center. A pilot study of 30 students has been conducted and we have data on a reading comprehension test taken before the program and a separate reading comprehension test taken after the program. Consider the data, found in the following Google Sheet.**

```{r, echo = TRUE}
# extract data
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1ijxX7c6qcFAw-r_mNSxC-Ps5b5njw-WtE4RcxPTySXU/edit?usp=sharing"))

# transform data to Reading dataframe
reading_data <- data %>% 
  # delete empty rows
  drop_na()

```

#### a. (From last week!) Use the appropriate *t*-test to determine if the program raises comprehension test averages by at least 5 points. Test at the &alpha;=0.05 level.

```{r, echo=TRUE}
# H0: Post <= Pre
# H1: Post > Pre
# p-Value < 0.001; Reject H0
t.test(data$Post, data$Pre,
       paired = TRUE,
       alternative = "greater")
```
### Hypothesis
H0: $\mu Post \le \mu Pre$\
H1: $\mu Post > \mu Pre$\

### Test
t0 = 4.9496

### p-Value
p < 0.001

## Conclusion
Reject H0 at the $\alpha$=0.05 level. There is sufficient evidence to suggest that the program raises comprehension test averages by at least 5 points.

#### b. Use the appropriate nonparametric test to determine if the program raises comprehension test averages by at least 5 points. Test at the &alpha;=0.05 level.

```{r, echo=TRUE}
wilcox.test(reading_data$Post, reading_data$Pre,
            paired = TRUE,
            alternative = "greater")
```
### Hypothesis
H0: $\mu Post \le \mu Pre$\
H1: $\mu Post > \mu Pre$\

### Test
t0 = 414

### p-Value
p < 0.001

## Conclusion
Reject H0 at the $\alpha$=0.05 level. There is sufficient evidence to suggest that the program raises comprehension test averages by at least 5 points.

#### c. Assess the assumption(s) associated with the *t*-test in part a. In your assessment, determine if the *t*-test or the corresponding nonparametric test should be reported.

### Assumption


```{r}
graph_paired_qq_plot <- function(data_table, after_sample, before_sample){
  data_table %>%
    mutate(diff = after_sample - before_sample) %>%
    ggplot(aes(sample = diff)) +
      stat_qq(size=3) +
      stat_qq_line() +
      theme_bw() +
      xlab("Theoretical") +
      ylab("Sample")
}
# Paired qq plot for Post - Pre reading scores
graph_paired_qq_plot(reading_data, reading_data$Post, reading_data$Pre)
```

### Frequency Reading
```{r}
# mean and median by dependent samples
# transform table from wide to long
data_long <- reading_data %>% 
  drop_na() %>% 
  pivot_longer(cols = Pre:Post, 
               names_to = "Test", 
               values_to = "Scores")
# summary statistics
summary_table <- data_long %>% 
  group_by(Test) %>%
  select(Test, Scores) %>% 
  summarise_all(list(Avg = mean,
                     Median = median))

change_table_format(summary_table)

# reading frequency
data_long %>% 
  ggplot(aes(x = Scores, colour = Test, fill = Test)) +
  geom_freqpoly(bins = 15)+
  theme_bw()

# reading density
data_long %>% 
  group_by(Test) %>% 
  ggplot(aes(x = Scores, color = Scores, fill = Test)) +
  geom_density(alpha=0.2)+
  geom_vline(data = summary_table, aes(xintercept=Avg, color=Test), linetype="dashed", size = 1.1)+
  xlim(35,110)+
  labs(title = "Test Scores", subtitle = "Average Increase")+
  theme_bw()+
  scale_fill_manual(values = custom_density_hue)+
  scale_color_manual(values = custom_density_hue)
```
## Assumption Conclusion
The T-test should be reported since the test scores are normal distributed. It is adequate to assume normality.
The density graph displays an improvement in the average scores.


#### d. Write a brief paragraph explaining your findings to the tutoring center, based on the test chosen in part c.
The tutoring department pilot study was a success. The reading comprehension should be integrated to their curriculum.
Moreover, record the scores for 3-6 months to have a bigger sample data.

### Function Tester 
```{r}
# bonus frequency
freq_test <- function(data, group_variable, num_variable){
  data %>% 
    ggplot(aes(x = num_variable, colour = group_variable, fill = group_variable)) +
    geom_freqpoly(bins = 50)+
    theme_bw()+
    scale_color_manual(values = custom_density_hue)
}

# bonus frequency
freq_test(bonus_data, bonus_data$Department, bonus_data$Bonus)
# reading frequency
freq_test(data_long, data_long$Test, data_long$Scores)
```