---
title: "Wilcoxons - Week 3, Lecture 1"
author: "Hector Gavilanes"
format: html
self-contained: true
editor: source
---

```{r setup, echo = TRUE, warnings = FALSE, message = FALSE}

library(tidyverse)
library(gsheet)
library(formattable)
library(gridExtra)
```

## Question 1:

**Recall the first dataset from last week. A large credit union awards quarterly bonuses to its employees. To ensure fairness across departments, data on the bonuses are regularly reviewed for equality. Consider the data in the following Google Sheet, which includes the bonus percentage for each employee in two departments.**

```{r, echo = TRUE}
# clear environment
rm(list = ls())

# extract data
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1q0Ouhr6hGBDp-5oyQuYGcsWTLrsayywVHJS7o17Sm6Y/edit?usp=sharing"))

# transform data to Bonus dataframe
bonus_data <- data %>% 
  # delete empty rows
  drop_na()

# ---CUSTOM COLORS---
custom_green_dark = "#71CA97"
custom_green_light = "#DeF7E9"
custom_gray = "#3D3D3D"
custom_orange = "#FC7307"
custom_purple = "#5A18C9"
custom_tan_light = "#F7E3D4"
custom_density_hue = c(custom_purple,custom_orange)
# ---FUNCTIONS---
# Change default table format
change_table_format <- function(data_table){
  # local variables
  first_column <- data_table[1]
  dataframe_length <- NCOL(data_table)
  # format
  formattable(data_table, align = c("l",rep("c", dataframe_length - 1)),
              list(first_column = formatter("span", style = ~ style(color = custom_gray, font.weight = "bold")),
              area(row = 1, col = 2:dataframe_length) ~ color_tile(custom_green_light, custom_green_dark)
              )#/end list
              )#/end formattable
}
```

#### a. (From last week!) Use the appropriate *t*-test to determine if the two departments are given different bonus amounts. Test at the &alpha;=0.01 level.

```{r, echo=TRUE}
# two-sided, two-sample t-test
t.test(bonus_data$Bonus ~ bonus_data$Department)
```

### Hypothesis
- $H_0$ : $\mu Credit = \mu Mortgage$
- $H_1$ : $\mu Credit \ne \mu Mortgage$

### Test Statistic
- $T_0 = 3.9$

### p-Value
- $p < 0.001$

## Conclusion
Reject $H_0$ at the $\alpha$=0.01 level. There is sufficient evidence to suggest that the two departments are given different bonuses.

#### b. Use the appropriate nonparametric test to determine if the two departments are given different bonus amounts. Test at the &alpha;=0.01 level.

```{r, echo=TRUE}
# nonparametric test for independent groups
wilcox.test(bonus_data$Bonus ~ bonus_data$Department,
            exact = FALSE)
```

### Hypothesis
- $H_0$ : $\ M credit = \ M mortgage$
- $H_1$ : $\ M credit \ne \ M mortgage$

### Test
- $T_0 = 683$

### p-Value
- $p < 0.001$

## Conclusion
Reject $H_0$ at the $\alpha$=0.01 level. There is sufficient evidence to suggest that the two departments are given different bonuses.

#### c. Assess the assumption(s) associated with the *t*-test in part a. In your assessment, determine if the *t*-test or the corresponding nonparametric test should be reported.

### Assumptions
```{r}
credit_qq <- bonus_data %>% 
  filter(Department == "Credit") %>% 
  ggplot(aes(sample = Bonus)) +
  stat_qq(size=3, alpha = 0.7, color = "magenta") +
  stat_qq_line() +
  theme_minimal() +
  xlab("Theoretical Credit") +
  ylab("Sample Credit")

mortgage_qq <- bonus_data %>% 
  filter(Department == "Mortgage") %>% 
  ggplot(aes(sample = Bonus)) +
    stat_qq(size=3, alpha = 0.6, color = custom_orange) +
    stat_qq_line() +
    theme_minimal() +
    xlab("Theoretical Mortgage") +
    ylab("Sample Mortgage")

# Combine both graphs
grid.arrange(credit_qq, mortgage_qq, nrow=1, bottom = "Bonus", top = "QQ Plot of Bonuses by Department")
```
### Assumption Supporting Graphs
```{r}
# mean by department
groups_mean <- bonus_data %>% 
  group_by(Department) %>% 
  summarise(Median = median(Bonus))
change_table_format(groups_mean)

# bonus density
bonus_data %>% 
  group_by(Department) %>% 
  ggplot(aes(x = Bonus, colour = Bonus, fill = Department)) +
  geom_density(alpha = 0.3)+
  geom_vline(data = groups_mean, aes(xintercept=Median, color=Department), linetype="dashed", size = 1.1)+
  xlim(5,13)+
  labs(title = "Bonuses by Department", subtitle = "Density Plot")+
  theme(legend.position="bottom")+
  theme_bw()+
  scale_fill_manual(values = custom_density_hue)+
  scale_color_manual(values = custom_density_hue)

```
The assumption of normality may not be valid; therefore, the report should use the Wilcoxon nonparametric test.
Furthermore, the QQ plot shows some discrepancies for the credit department, with some outliers towards the right tail. The density graph displays the median for both departments. 

#### d. Write a brief paragraph explaining your findings to management, based on the test chosen in part c.

## Conclusion
The mortgage department needs improvement. Perhaps the credit department can share insights into business processes to help the mortgage department. The mortgage department needs further investigation.


## Question 2: 

**Recall the second dataset from last week. A new 2 hour reading comprehension program has been developed by a local tutoring center. A pilot study of 30 students has been conducted and we have data on a reading comprehension test taken before the program and a separate reading comprehension test taken after the program. Consider the data, found in the following Google Sheet.**

```{r, echo = TRUE}
# extract data
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1ijxX7c6qcFAw-r_mNSxC-Ps5b5njw-WtE4RcxPTySXU/edit?usp=sharing"))

# transform data to Reading dataframe
reading_data <- data %>% 
  # delete empty rows
  drop_na()

```

#### a. (From last week!) Use the appropriate *t*-test to determine if the program raises comprehension test averages by at least 5 points. Test at the &alpha;=0.05 level.

```{r, echo=TRUE}
# H0: Post - Pre <= 5
# H1: Post - Pre > 5
# mu = 5
t.test(data$Post, data$Pre,
       paired = TRUE,
       mu = 5,
       alternative = "greater")
```
### Hypothesis
- $H_0$ : $\mu Post - \mu Pre \le 5$\
- $H_1$ : $\mu Post - \mu Pre > 5$\

### Test Statistic
- $T_0 = -1.56$

### p-Value
- $p = 0.94$

## Conclusion
Keep $H_0$ at the $\alpha$=0.05 level. There is not sufficient evidence to suggest that the program raises comprehension test averages by at least 5 points.

#### b. Use the appropriate nonparametric test to determine if the program raises comprehension test averages by at least 5 points. Test at the &alpha;=0.05 level.

```{r, echo=TRUE}
# nonparametric test for dependent groups
wilcox.test(reading_data$Post, reading_data$Pre,
            paired = TRUE,
            mu = 5,
            alternative = "greater",
            exact = FALSE)
```
### Hypothesis
- $H_0$ : $\ M post - \ M pre \le 5$
- $H_1$ : $\ M post - \ M pre > 5$

### Test Statistic
- $T_0 = 124.5$

### p-Value
- $p = 0.91$

## Conclusion
Keep $H_0$ at the $\alpha$=0.05 level. There is not sufficient evidence to suggest that the program raises comprehension test averages by at least 5 points.

#### c. Assess the assumption(s) associated with the *t*-test in part a. In your assessment, determine if the *t*-test or the corresponding nonparametric test should be reported.

### Assumption

```{r}
graph_paired_qq_plot <- function(data_table, after_sample, before_sample){
  data_table %>%
    mutate(diff = after_sample - before_sample) %>%
    ggplot(aes(sample = diff)) +
    stat_qq(size=3, alpha = 0.6, color = custom_orange) +
    stat_qq_line() +
    theme_bw() +
    labs(title = "Reading Comprehension Test", subtitle = "QQ Plot")+
    xlab("Theoretical Reading") +
    ylab("Sample Reading")
}
# Paired qq plot for Post - Pre reading scores
graph_paired_qq_plot(reading_data, reading_data$Post, reading_data$Pre)
```

### Assumption Supporting Graphs
```{r}
# transform table from wide to long
data_long <- reading_data %>% 
  drop_na() %>% 
  pivot_longer(cols = Pre:Post, 
               names_to = "Test", 
               values_to = "Scores")
# summary statistics
# median by dependent samples
summary_long_table <- data_long %>% 
  group_by(Test) %>%
  select(Test, Scores) %>% 
  summarise_all(list(Median = median))

change_table_format(summary_long_table)

# reading density
data_long %>% 
  group_by(Test) %>% 
  ggplot(aes(x = Scores, color = Scores, fill = Test)) +
  geom_density(alpha=0.3)+
  geom_vline(data = summary_long_table, aes(xintercept=Median, color=Test), linetype="dashed", size = 1.1)+
  xlim(35,110)+
  labs(title = "Reading Comprehension Test", subtitle = "Density Plot")+
  theme_bw()+
  scale_fill_manual(values = custom_density_hue)+
  scale_color_manual(values = custom_density_hue)
```
## Assumption Conclusion
It is adequate to assume normality; thus, the report should use the T-test.

#### d. Write a brief paragraph explaining your findings to the tutoring center, based on the test chosen in part c.
Even though the scores have improved, the grade average is not greater than 5 points. The tutoring department should continue working on the reading comprehension program. Yet, the reading comprehension program is not ready to be integrated into its curriculum. Moreover, a bigger sample size will help helpful.