---
title: "Two-Sample Means (Dependent) - Week 2, Lecture 2"
author: "Hector Gavilanes"
format: html
self-contained: true
editor: source
---

```{r setup, echo = TRUE, warnings = FALSE, message = FALSE}
#install.packages("formattable")
library(tidyverse)
library(gsheet)
library(formattable)
library(gridExtra)
```

**A new 2 hour reading comprehension program has been developed by a local tutoring center. A pilot study of 30 students has been conducted and we have data on a reading comprehension test taken before the program and a separate reading comprehension test taken after the program. Consider the data, found in the following Google Sheet.**

```{r, echo = TRUE}
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1ijxX7c6qcFAw-r_mNSxC-Ps5b5njw-WtE4RcxPTySXU/edit?usp=sharing"))

# ---FUNCTIONS---
# Change default table format
change_table_format <- function(data_table){
  formattable(data_table, align =c("l",rep("c", NCOL(data_table) - 1)))
}
custom_green_dark = "#DeF7E9"
custom_green_light = "#71CA97"
custom_red = "#ff7f7f"

```

#### 1. Summarize the data. Provide appropriate summary statistics for the pilot study.

```{r, echo=TRUE}
# transform table from wide to long
data_long <- data %>% 
  pivot_longer(cols = Pre:Post, 
               names_to = "Test", 
               values_to = "Scores")
# summary statistics
summary_table <- data_long %>% 
  group_by(Test) %>%
  select(Test, Scores) %>% 
  summarise_all(list(Avg = mean,SD = sd, Median = median, Min = min, Max = max))
# change table format
change_table_format(summary_table)
```

#### 2. Construct the 95% confidence interval for the difference in test scores. Hint: we want to discuss this in terms of the *increase* in test scores.

```{r, echo=TRUE}
t.test(data$Post, data$Pre,
       paired = TRUE)
```

The 95% CI are (2.23, 5.37)

#### 3. Use the appropriate hypothesis test to determine if the program raises comprehension test averages by at least 5 points. Test at the &alpha;=0.05 level.

```{r, echo=TRUE}

# H0: Post <= Pre
# H1: Post > Pre
# p-Value < 0.001; Reject H0
t.test(data$Post, data$Pre,
       paired = TRUE,
       alternative = "greater")

```
### Hypothesis
H0: $\mu_Post \le \mu_Pre$\
H1: $\mu_Post > \mu_Pre$\

### Test
t0 = 4.9496

### p-Value
p < 0.001

## Conclusion
Reject H0 at the $\alpha$=0.05 level. There is sufficient evidence to suggest that the program raises comprehension test averages by at least 5 points.


#### 4. (Challenge!) Construct a graph that will help you explain your findings to the tutoring center.

### Boxplots - Reading Pilot Program
```{r}
# Graph Boxplot
plot_boxes <- function(y_cord, fill_color, x_label, txt_color){
  data %>%
    ggplot(aes(y=y_cord))+
    geom_boxplot(fill=fill_color, alpha=0.4, size=1.2, notch = TRUE, notchwidth = 0.6)+
    xlab(x_label)+
    ylab("")+
    ylim(40,100)+
    theme_minimal()+
    theme(text=element_text(color=txt_color))
}

pre_plot <- plot_boxes(data$Pre, "magenta", "PRE", "purple")
post_plot <- plot_boxes(data$Post, "cyan", "POST", "#009999")
grid.arrange(pre_plot, post_plot, nrow=1, bottom = "Reading Program", left = "Scores", top = "Reading Comprehension Program")
```



#### 5. Write a brief paragraph explaining your findings to the tutoring center. 

The tutoring center should continue developing similar programs for other classes because students are improving their grade scores. 







