---
title: "ANOVA Assumptions and the Kruskal-Wallis - Week 5, Lecture 2"
author: "Hector Gavilanes"
format: html
self-contained: true
editor: source
---

```{r setup, echo = TRUE, warnings = FALSE, message = FALSE}
#install.packages(DescTools)
library(tidyverse)
library(gsheet)
library(formattable)
library(DescTools) # perform Dunnett Test
```

**Recall the agricultural data from the ANOVA activity available in the following [Google Sheet](https://docs.google.com/spreadsheets/d/1CFyuuWy-4YmVU3wkYuC8Uf3xW3qkpCnpF468NjBAhbY/edit#gid=0).**

```{r}
# clear environment
rm(list = ls())

# extract data
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1CFyuuWy-4YmVU3wkYuC8Uf3xW3qkpCnpF468NjBAhbY/edit#gid=0"))

# transform data
clean_data <- data %>% 
  # delete empty rows
  drop_na()
# pivot to long data
long_data <- clean_data %>% 
  pivot_longer(cols = c("Herbicide1", "Herbicide2", "Control"),
               names_to = "Herbicide_groups",
               values_to = "Results")

# ---Colors---
custom_green = "#7FBD32"
custom_purple = "#700CBC"
custom_orange = "#FB8604"

# ***FUNCTIONS***
# f01 // change default table format
change_table_format <- function(table_name){
  formattable(
  table_name, align =c("c","c", "c"),
  list(area() ~ color_text("black", "#555888"))
  )
}
```

#### 1. State the ANOVA assumptions mathematically.
```{r}
# one-way ANOVA Hypothesis test
module_data_anova <- aov(long_data$Results ~ long_data$Herbicide_groups)
summary(module_data_anova)
```
### Hypothesis
- $H_0$: $\mu Control = \mu HerbicideA\ = \mu HerbicideB$\
- $H_1$: at least one is different

### Test Statistic
- $F_0 = 3.677$

### p-Value
- $p = 0.0293$

## Conclusion
Reject $H_0$ at the $\alpha$=0.05 level. There is sufficient evidence to suggest that there is a difference in the average yield between the three groups.

#### 2. Explain how to check the ANOVA assumptions.

The one-way ANOVA hypothesis test indicates a difference between the groups. However, there is a need to determine which groups are different. Moreover, ANOVA assumptions do not check for the normal distribution of the data. 

In this case, QQ plots and histograms help to check visually for normality of residuals. 

It is good practice to check the normal distribution from the continuous variable because of the chances of having false positive results increase. Thus, a formal statistical test like Kruskal-Wallis is a nonparametric alternative to one-way ANOVA to check for normality.

#### 3. Assess the ANOVA assumptions for the agricultural data, graphically. In addition to providing the relevant graphs, you must provide an assessment for each assumption.

```{r, echo=TRUE}
plot(module_data_anova)
```

```{r}
# histogram
hist(resid(module_data_anova))
```


The QQ
#### 4. Use the appropriate hypothesis test to assess the variance assumption. Test at the $\alpha=0.05$ level.

```{r, echo=TRUE}

```

#### 5. Provide an overall assessment -- do we meet the assumptions to use ANOVA?

#### 6. Perform the Kruskal-Wallis to determine if there is a difference between the treatment groups. Test at the $\alpha=0.05$ level.

```{r, echo=TRUE}

```

#### 7. Perform the Kruskal-Wallis post-hoc test to determine what differences exist between groups. Test at the $\alpha=0.05$ level.

#### 8. Challenge! Construct a graph (or multiple graphs) to aid in explaining this to the agricultural researcher.

#### 9. Write a summary paragraph, appropriate for the agricultural researcher (i.e., not a statistician), using the results that are **appropriate for the data**. Hint: you will summarize either your ANOVA findings from the previous activity *or* your Kruskal-Wallis findings from this activity.

