---
title: "ANOVA Assumptions and the Kruskal-Wallis - Week 5, Lecture 2"
author: "Hector Gavilanes"
format: html
self-contained: true
editor: source
---

```{r setup, echo = TRUE, warnings = FALSE, message = FALSE}
# install.packages("DescTools")
# install.packages("pgirmess")
# install.packages("ggstatsplot")
# install.packages("ISLR")
# install.packages("PMCMRplus")
library(tidyverse)
library(gsheet)
library(formattable)
library(DescTools) # perform Dunnett, Levene Test
library(pgirmess) # perform Kruskal-Wallis Post-hoc
library(ggstatsplot)
library(ISLR)
library(patchwork)
library(PMCMRplus)
```

**Recall the agricultural data from the ANOVA activity available in the following [Google Sheet](https://docs.google.com/spreadsheets/d/1CFyuuWy-4YmVU3wkYuC8Uf3xW3qkpCnpF468NjBAhbY/edit#gid=0).**

```{r}
# clear environment
rm(list = ls())

# extract data
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1CFyuuWy-4YmVU3wkYuC8Uf3xW3qkpCnpF468NjBAhbY/edit#gid=0"))

# transform data
clean_data <- data %>% 
  # delete empty rows
  drop_na()
# pivot to long data
long_data <- clean_data %>% 
  pivot_longer(cols = c("Herbicide1", "Herbicide2", "Control"),
               names_to = "Herbicide_groups",
               values_to = "Results")

# ---Colors---
custom_gray = "#3D3D3D"
custom_green_dark = "#71CA97"
custom_green_light = "#DeF7E9"
custom_green = "#7FBD32"
custom_purple = "#700CBC"
custom_orange = "#FB8604"
custom_density_hue = c(custom_purple,custom_orange, custom_green)

# ***FUNCTIONS***
# f00 - Change default table format
change_table_format <- function(data_table){
  # local variables
  first_column <- data_table[1]
  dataframe_length <- NCOL(data_table)
  # format
  formattable(data_table, align = c("l",rep("c", dataframe_length - 1)),
              list(first_column = formatter("span", style = ~ style(color = custom_gray, font.weight = "bold"))
              )#/end list
              )#/end formattable
}
```

#### 1. State the ANOVA assumptions mathematically.
```{r}
# ANOVA table
module_data_anova <- aov(long_data$Results ~ long_data$Herbicide_groups)
summary(module_data_anova)

# Summary table
summary_table <- long_data %>% 
  drop_na() %>% 
  group_by(Herbicide_groups) %>% 
  summarise(SD = sd(Results), AVG = mean(Results), Median = median(Results)) %>% 
  ungroup()

change_table_format(summary_table)

# Levene's test for equality of variance, 
# default: median, provides a more robust test (Brown-Forsythe-Test)
LeveneTest(long_data$Results ~ long_data$Herbicide_groups)
```
The data failed the Levene's Test with a $p-value = 0.04688$. 

#### 2. Explain how to check the ANOVA assumptions.

The one-way ANOVA hypothesis test indicates a difference between the groups. However, there is a need to determine which groups are different. Moreover, a residual analysis will check for the normal distribution of the data. 

In this case, residual plots can be generated with QQ plots, or histograms to check visually for normality of residuals. 

It is good practice to check for normality from the continuous variable because of the chances of having false positive results increase. If the assumption fails, a formal statistical test like Kruskal-Wallis is a nonparametric alternative to one-way ANOVA.

The one-way ANOVA assumptions are listed below.

Normality: The variable under discussion is regularly distributed for each population.

Sample independence: Each sample is drawn separately from the others.

Variance equality: The data variance in each group is the same.

Simple random samples: Simple random samples are samples taken from the populations under study.

#### 3. Assess the ANOVA assumptions for the agricultural data, graphically. In addition to providing the relevant graphs, you must provide an assessment for each assumption.

```{r, echo=TRUE}
plot(module_data_anova)
```
The QQ plot displays normality of residuals. There are a few outliers towards the right tail. Nonetheless, the outliers do not represent a concern.

```{r}
# histogram
hist(resid(module_data_anova))
```

The histogram is positively skewed; however, the data is relatively normal. Thus, the graph does not represent a cause for concern.

#### 4. Use the appropriate hypothesis test to assess the variance assumption. Test at the $\alpha=0.05$ level.

```{r}
# one-way ANOVA Hypothesis test
module_data_anova <- aov(long_data$Results ~ long_data$Herbicide_groups)
summary(module_data_anova)
```
### Hypothesis
- $H_0$: $\mu Control = \mu HerbicideA\ = \mu HerbicideB$\
- $H_1$: at least one group is different

### Test Statistic
- $F_0 = 3.677$

### p-Value
- $p = 0.0293$

## Conclusion
Reject $H_0$ at the $\alpha$=0.05 level. There is sufficient evidence to suggest that there is a difference in the average yield between the three groups.

#### 5. Provide an overall assessment -- do we meet the assumptions to use ANOVA?
Since there are no significant outliers, and the data is approximately normally distributed, it is adequate to use ANOVA.
The data failed the Levene's Test with a $p-value = 0.04688$; however, I feel confident with the ANOVA assumptions.

#### 6. Perform the Kruskal-Wallis to determine if there is a difference between the treatment groups. Test at the $\alpha=0.05$ level.

```{r, echo=TRUE}
kruskal.test(long_data$Results ~ long_data$Herbicide_groups)
```

### Hypothesis: Kruskal-Wallis
- $H_0$: $M_{Control} = M_A = M_B$\
- $H_1$: at least one group is different

### Test Statistic
- $H = 8.6303$

### p-Value
- $p = 0.01336$

## Conclusion
Reject $H_0$ at the $\alpha$=0.05 level. There is sufficient evidence to suggest that there is a difference in the average yield between the three groups.

#### 7. Perform the Kruskal-Wallis post-hoc test to determine what differences exist between groups. Test at the $\alpha=0.05$ level.
```{r}
# Kruskal-Wallis post-hoc
kruskalmc(long_data$Results ~ long_data$Herbicide_groups)

# Multiple pairwise-comparison between groups
pairwise.wilcox.test(long_data$Results, long_data$Herbicide_groups,
                     p.adjust.method = "bonferroni",
                     exact = FALSE)
```
- Only Herbicide A and B are significant different with a $p-value = 0.015$.

ANOVA and Kruskal-Wallis tests led to the same conclusion.

#### 8. Challenge! Construct a graph (or multiple graphs) to aid in explaining this to the agricultural researcher.

The following QQ plot displays how the values lie on the 45-degree line relatively well. The data is approximately normally distributed.
```{r}
# QQ plot
ggplot(module_data_anova, aes(sample = module_data_anova$residuals))+
  geom_qq(size=3, alpha = 0.6, color = custom_orange)+
  geom_qq_line()+
  theme_bw() +
  labs(title = "Normal Q-Q Plot")+
  xlab("Theoretical Quantiles") +
  ylab("Standarized Residuals")
```

The histogram quickly matches the general appearance  of a normal distribution.
The data is skewed; nonetheless, it does not represent a cause for concern.
```{r}
# histogram
ggplot(lm(long_data$Results ~ long_data$Herbicide_groups)) + 
  geom_histogram(aes(x=module_data_anova$residuals), 
                 binwidth=3, fill=custom_green, color="#e9ecef", alpha=0.9)+
  xlab("Standarized Residuals")+
  ylab("Frequency")+
  theme_bw()
```
The graph below shows the fitted vs residuals.
The plot is used to detect non-linearity, unequal error variances, and outliers.
```{r}
# fitted vs residuals
ggplot(lm(long_data$Results ~ long_data$Herbicide_groups)) + 
  geom_point(aes(x=.fitted, y=.resid), 
             size=3, alpha = 0.35, color = custom_purple)+
  xlab("Fitted")+
  ylab("Standarized Residuals")+
  theme_bw()
```

The following density graph displays the variance among the three groups.
```{r}
# density plot
density_plot <- long_data %>%  
  ggplot(aes(Results, fill = factor(Herbicide_groups)))+
  geom_density(alpha = 0.6)+
  xlab("Herbicide Groups")+
  theme_bw()+
  scale_fill_manual(values = custom_density_hue)+
  scale_color_manual(values = custom_density_hue)+
  geom_vline(xintercept = summary_table$AVG, size = 1.5, color = custom_purple, alpha(0.55))+
  geom_vline(xintercept = summary_table$Median, size = 1, color = "red", linetype="dashed")
density_plot
```
The following chart displays all three groups with a significant difference outlined between Herbicide A and B.
The Bonferroni correction is used to reduce the chances of obtaining false-positive results.
```{r}
ggbetweenstats(
  data = long_data,
  x = Herbicide_groups,
  y = Results,
  plot.type = "box",
  type = "p",
  pairwise.display = "s",
  var.equal = T,
  p.adjust.method = "bonferroni",
  ylab = "Crop Yield of Tomatoes",
  xlab = "Herbicide Groups",
  k = 0,
  bf.message = F
)
```

#### 9. Write a summary paragraph, appropriate for the agricultural researcher (i.e., not a statistician), using the results that are **appropriate for the data**. Hint: you will summarize either your ANOVA findings from the previous activity *or* your Kruskal-Wallis findings from this activity.

Since there are no significant outliers, and the data is approximately normally distributed, it is reasonable to make the ANOVA assumptions.
Our findings showed that there is no significant difference against the control group.
However, there is a difference between Herbicide A and B. 

To conclude, Herbicide B showed consistent and better performance on crop yields of tomatoes against the Tomato Hornworms.