---
title: "Logistic Regression - Week 12, Day 1"
author: "Hector Gavilanes"
format: html
self-contained: true
editor: source
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(gsheet)
library(naniar)
```

**Recall the Titanic data from lecture. Use it to answer the following questions.**
```{r}
# clear environment
rm(list = ls())
# load data
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1TrJ6jZ3hP6gQk6FvqbMrFUysEUEqle7XfuLnAkZsHYI/edit"))
```

**1. Model survival as a function of age (*Age*), the number of siblings and/or spouses (*SibSp*), the number of parents or children (*Parch*), the class of ticket (*Pclass*), and amount paid for ticket (*Fare*).** 

```{r, echo = TRUE}
m1 <- glm(data$Survived ~ data$Age + data$SibSp + data$Parch + data$Pclass + data$Fare,
          family = binomial,
          na.action = na.exclude)
summary(m1)
```

**1a. Report the model.**
```{r}
m1_coeff <- coefficients(m1)
```

$$
\ln \left( \frac{\hat{\pi}}{1-\hat{\pi}} \right) = `r round(m1_coeff[1],3)` `r round(m1_coeff[2],3)` \text{ Age} `r round(m1_coeff[3],3)` \text{ Sibling \& Spouse} + `r round(m1_coeff[4],3)` \text{ Parents \& Children} `r round(m1_coeff[5],3)` \text{ Class} + `r round(m1_coeff[6],3)` \text{ Fare}
$$
**1b. Which, if any, are significant predictors of survival? The full hypothesis test is not needed, but you must provide statistical justification.**

```{r, echo = TRUE}
significant <- coef(summary(m1))[,'Pr(>|z|)']
significant < 0.05
round(significant, 3)

non_significant <- coef(summary(m1))[,'Pr(>|z|)']
non_significant > 0.05
round(non_significant, 3)
```

- Age (p < 0.001), Siblings & Spouses (p = 0.006), Parents & Children (p = 0.023), and Passenger Class (p < 0.001) are significant predictors of survival.
- Fare (0.194) is not a significant predictor of survival. 


**1c. Find the odds ratios and corresponding 95% confidence intervals.**

```{r, echo = TRUE}
# get odd ratios
round(exp(coefficients(m1)[-1]), 3)

# get CI of odd ratios
round(exp(confint(m1)), 3)

```



**1d. Provide interpretations for the odds ratios for each of the predictors in the model.**

- For a 1 year increase in age, the odds of a traveler surviving is multiplied by 0.956, which is a 4.4% decrease. 
- For a 1 person increase in sibling/spouse, the odds of a traveler surviving is multiplied by 0.747, which is a 25.3% decrease. 
- For a 1 person increase in parents/children, the odds of a traveler surviving is multiplied by 1.281, which is a 28.1% increase.
- For a 1 unit increase in class (going to a different level), the odds of a traveler surviving is multiplied by 0.316, which is a 68.4% decrease.
- For a $1 increase in fare paid, the odds of a traveler surviving is multiplied by 1.003, which is a 0.3% increase.

**2. Challenge! Model survival as a function of age (*Age*), the class of ticket (*Pclass*), and sex of passenger (*Sex*).** 

```{r, echo = TRUE}
# change variables to a factor (categorical)
data$Sex <- as.factor(data$Sex)
data$Embarked <- as.factor(data$Embarked)

# check for missing values
miss_var_summary(data)
# check for unique values
sapply(data, function(x) length(unique(x)))

# visualizing missing values
gg_miss_var(data) +
  labs(title = "Missing Values vs Observed",
       y = "Missing Values")+
  guides(colour = "none")
 
# drop Cabin, Passenger-ID, Name, Ticket columns
clean_data <- subset(data,select=c(2,3,5,6,7,8,10,12))

# replace missing values with mean for Age
clean_data$Age[is.na(clean_data$Age)] <- mean(clean_data$Age,na.rm=T)

# double-check if variables are factors 
is.factor(clean_data$Sex)
is.factor(clean_data$Embarked)

# R interpretation of categorical variable
contrasts(clean_data$Sex)

# logistic regression model, excluding missing values
m2 <- glm(Survived ~ Age + Pclass + Sex,
          data = clean_data,
          family = binomial,
          na.action = na.exclude)
summary(m2)

```

**2a. Report the model.**
```{r}
m2_coeff <- coefficients(m2)
m2_coeff
```

$$
\ln \left( \frac{\hat{\pi}}{1-\hat{\pi}} \right) = `r round(m2_coeff[1],3)` `r round(m2_coeff[2],3)` \text{ Age} `r round(m2_coeff[3],3)` \text{ Class} `r round(m2_coeff[4],3)` \text{ Sex} 
$$
**2b. Which, if any, are significant predictors of survival? The full hypothesis test is not needed, but you must provide statistical justification.**

```{r, echo = TRUE}
significant <- coef(summary(m2))[,'Pr(>|z|)']
significant < 0.05
round(significant, 3)

non_significant <- coef(summary(m2))[,'Pr(>|z|)']
non_significant > 0.05
round(non_significant, 3)
```
All variables are significant predictors of survival:
- Age     (p = 0.004)
- Class   (p < 0.001)
- Sex     (p < 0.001)

**2c. Find the odds ratios and corresponding 95% confidence intervals.**

```{r, echo = TRUE}
# get odds ratios
round(exp(coefficients(m2)[-1]), 3)

# get CI of odd ratios
exp(confint(m2))

```

**2d. Provide interpretations for the odds ratios for each of the predictors in the model.**

- For a 1 year increase in age, the odds of a traveler surviving is multiplied by 0.967, which is a 3.3% decrease. 
- For a 1 unit increase in class (going to a different level), the odds of a traveler surviving is multiplied by 0.311, which is a 68.9% decrease.
- For a 1 male increase in sex, the odds of a traveler surviving is multiplied by 0.073, which is a 92.7% decrease.
