---
title: "Two-Sample Means (Dependent) - Week 2, Lecture 2"
author: "Hector Gavilanes"
format: html
self-contained: true
editor: source
---

```{r setup, echo = TRUE, warnings = FALSE, message = FALSE}
#install.packages("formattable")
library(tidyverse)
library(gsheet)
library(formattable)
library(gridExtra)
```

**A new 2 hour reading comprehension program has been developed by a local tutoring center. A pilot study of 30 students has been conducted and we have data on a reading comprehension test taken before the program and a separate reading comprehension test taken after the program. Consider the data, found in the following Google Sheet.**

```{r, echo = TRUE}
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1ijxX7c6qcFAw-r_mNSxC-Ps5b5njw-WtE4RcxPTySXU/edit?usp=sharing"))

# ---FUNCTIONS---
# Change default table format
change_table_format <- function(data_table){
  formattable(data_table, align =c("l",rep("c", NCOL(data_table) - 1)))
}
```

#### 1. Summarize the data. Provide appropriate summary statistics for the pilot study.

```{r, echo=TRUE}
# summary with standard deviation by department
summary_table <- data %>% 
  drop_na() %>%
  select(Pre, Post) %>% 
  summarise_all(list(Avg = mean, 
                      SD = sd, 
                      Median = median,
                      Variance = var,
                      Min = min,
                      Max = max))

change_table_format(summary_table)



```

#### 2. Construct the 95% confidence interval for the difference in test scores. Hint: we want to discuss this in terms of the *increase* in test scores.

```{r, echo=TRUE}
t.test(data$Pre, data$Post,
       paired = TRUE)
```

The 95% CI is (-5.37, -2.23)

#### 3. Use the appropriate hypothesis test to determine if the program raises comprehension test averages by at least 5 points. Test at the &alpha;=0.05 level.

```{r, echo=TRUE}
t.test(data$Pre, data$Post,
       paired = TRUE,
       alternative = "greater")
```
### Hypothesis
H0: $\mu_Pre \le \mu_Post$\
H1: $\mu_Pre > \mu_Post$\

### Test
t0 = -11.46

### p-Value
p = 1

## Conclusion
Keep H0 at the $\alpha$=0.05 level. There is sufficient evidence to suggest that the program raises comprehension test averages by at least 5 points.


#### 4. (Challenge!) Construct a graph that will help you explain your findings to the tutoring center.

### Assumptions - Reading Pilot Program
```{r, echo=TRUE}
qq_graph <- data %>%
  mutate(diff = Pre - Post) %>%
  ggplot(aes(sample = diff))+
  stat_qq(size = 3)+
  stat_qq_line()+
  theme_bw()+
  xlab("Theoretical")+
  ylab("Sample")
qq_graph
```

### Boxplots - Reading Pilot Program
```{r}
# Graph Boxplot
plot_boxes <- function(y_cord, fill_color, x_label, txt_color){
  data %>%
    ggplot(aes(y=y_cord))+
    geom_boxplot(fill=fill_color, alpha=0.4, size=1.2, notch = TRUE, notchwidth = 0.6)+
    xlab(x_label)+
    ylab("")+
    ylim(40,100)+
    theme_minimal()+
    theme(text=element_text(color=txt_color))
}

pre_plot <- plot_boxes(data$Pre, "magenta", "PRE", "purple")
post_plot <- plot_boxes(data$Post, "cyan", "POST", "#009999")
grid.arrange(pre_plot, post_plot, nrow=1, bottom = "Reading Program", left = "Scores", top = "Reading Comprehension Program")
```



#### 5. Write a brief paragraph explaining your findings to the tutoring center. 

The tutoring center should continue developing similar programs for other classes because students are improving their grade scores. 







