---
title: "One-Way ANOVA - Week 5, Lecture 1"
author: "Hector Gavilanes"
format: html
self-contained: true
editor: source
---

```{r setup, echo = TRUE, warnings = FALSE, message = FALSE}
#install.packages(DescTools)
library(tidyverse)
library(gsheet)
library(formattable)
library(DescTools) # perform Dunnett Test
```

**An agricultural researcher investigated two new herbicides and their effects on crop yields of tomatoes, which can be demolished by the [tomato hornworm](https://extension.umn.edu/yard-and-garden-insects/tomato-hornworms) in mere hours. There were three areas for the experiment: an area for herbicide A, an area for herbicide B, and an area for no herbicide at all (control). At the end of the growing season, the yields (in bushels per acre) were measured and are available in the following [Google Sheet](https://docs.google.com/spreadsheets/d/1CFyuuWy-4YmVU3wkYuC8Uf3xW3qkpCnpF468NjBAhbY/edit#gid=0).**

```{r, echo=TRUE}
# clear environment
rm(list = ls())

# extract data
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1CFyuuWy-4YmVU3wkYuC8Uf3xW3qkpCnpF468NjBAhbY/edit#gid=0"))

# transform data
clean_data <- data %>% 
  # delete empty rows
  drop_na()
# pivot to long data
long_data <- clean_data %>% 
  pivot_longer(cols = c("Herbicide1", "Herbicide2", "Control"),
               names_to = "Herbicide_groups",
               values_to = "Results")

# ---Colors---
custom_green = "#7FBD32"
custom_purple = "#700CBC"
custom_orange = "#FB8604"

# ***FUNCTIONS***
# f01 // change default table format
change_table_format <- function(table_name){
  formattable(
  table_name, align =c("c","c", "c"),
  list(area() ~ color_text("black", "#555888"))
  )
}
# f02 // get confidence intervals for each group
get_ci <- function(df, category, continue_variable){
  ci <- df %>% 
    group_by(!! sym(category)) %>%
    summarise(
      ci_lower = t.test(!! sym(continue_variable), conf.level = 0.95)$conf.int[1],
      ci_upper = t.test(!! sym(continue_variable), conf.level = 0.95)$conf.int[2]
    )
  return(ci)
}
# f03 // get mean, confidence intervals for each group
get_mean_ci <- function(df, category, continue_variable){
  ci <- df %>% 
    group_by(!! sym(category)) %>%
    summarise(
      mean = mean(!! sym(continue_variable)),
      ci_lower = t.test(!! sym(continue_variable), conf.level = 0.95)$conf.int[1],
      ci_upper = t.test(!! sym(continue_variable), conf.level = 0.95)$conf.int[2]
    )
  return(ci)
}
```

#### 1. Construct the 95% confidence interval for the average yield for each group (A, B, control).

```{r, echo=TRUE}
# print CI table for the average yield for each group
ci <- get_ci(long_data, 'Herbicide_groups', 'Results')
change_table_format(ci)
```

#### 2. Challenge! Construct a graph to display all confidence intervals.

```{r, echo=TRUE}
# get mean, CI for each group
mean_ci <- get_mean_ci(long_data, 'Herbicide_groups', 'Results')

# Plot Confidence Intervals
ggplot(mean_ci, aes(x = Herbicide_groups, y = mean))+
  # geom_line(aes(x = Herbicide_groups, y = mean))+
  geom_point(aes(x = Herbicide_groups, y = mean), color = "black", size = 3)+
  scale_color_manual(name = "Herbicide Groups", values=c(custom_green, custom_purple, custom_orange))+
  geom_errorbar(aes(x = Herbicide_groups, ymin = ci_lower, ymax = ci_upper, 
                    color = factor(Herbicide_groups)), size = 1.5, width = 0.4)+
  geom_text(aes(y = ci_lower, label = round(ci_lower,1)), size = 4, vjust = 2, color = "black")+
  geom_text(aes(y = ci_upper, label = round(ci_upper,1)), size = 4, vjust = -1, color = "black")+
  labs(title = "Chart with 95% Confidence Intervals")+
  labs(x="Herbicide Groups", y="Yield Results")+
  theme(legend.position="none")+
  ylim(80, 95)+
  theme_bw()
```

#### 3. Challenge! Why should we be cautious when using the individual CIs found in Q1 and displayed in Q2 to directly compare the three groups?
Visually, the CI overlapping method fails to reject the null hypothesis more frequently than the corresponding hypothesis test. Using this method decreases the statistical power of an adequate assessment, potentially causing to miss differences.

#### 4. Use the appropriate hypothesis test to show that there is a difference in the average yield between the three groups. Test at the $\alpha=0.05$ level.

```{r, echo=TRUE}
# one-way ANOVA Hypothesis test
module_data_anova <- aov(long_data$Results ~ long_data$Herbicide_groups)
summary(module_data_anova)
```
### Hypothesis
- $H_0$: $\mu Control = \mu HerbicideA\ = \mu HerbicideB$\
- $H_1$: at least one is different

### Test Statistic
- $F_0 = 3.677$

### p-Value
- $p = 0.0293$

## Conclusion
Reject $H_0$ at the $\alpha$=0.05 level. There is sufficient evidence to suggest that there is a difference in the average yield between the three groups.

#### 5. Construct and test the following contrasts for the following comparisons:

##### a. Herbicide A vs. Herbicide B. Test at the $\alpha=0.05$ level.

```{r, echo=TRUE}
# Post-hoc Testing: Linear Contrasts
# check the order of the groups
long_data$Herbicide_groups <- as.factor(long_data$Herbicide_groups)
contrasts(long_data$Herbicide_groups)
# adjust the alpha level
# assigning specific values to the matrix depending on grouping
contrasts(long_data$Herbicide_groups) <- cbind(c(0, -1, 1), 
                                               c(-1, 0.5, 0.5))
```
After the Order Verification Process, proceed with the specific linear contrast.

```{r, echo=TRUE}
# ANOVA test: A vs B
m1 <- aov(Results ~ Herbicide_groups, 
          data = long_data)
summary.aov(m1, split=list(Herbicide_groups=list("A vs. B"=1)))
```
There is a difference between Herbicide A and B (p = 0.00825).

Paiwise t-test with Bonferroni Correction
```{r}
pairwise.t.test(long_data$Results, long_data$Herbicide_groups, p.adjust.method = "bonferroni")
```
The only significant difference with a p-value = 0.025 is between Herbicide A vs Herbicide B .

### Hypothesis - Linear Contrasts
- $H_0$: $\mu HerbicideA\ = \mu HerbicideB$\
- $H_1$: $\mu HerbicideA\ \ne  \mu HerbicideB$\

### Test Statistic
- $F_0 = 7.310$

### p-Value
- $p = 0.00825$

## Conclusion
Reject $H_0$ at the $\alpha$=0.05 ($\alpha_B$ = 0.025) level. There is sufficient evidence to suggest that there is a difference in the average yield between the three groups.

##### b. Herbicide A and Herbicide B vs. Control. Test at the $\alpha=0.05$ level.

```{r, echo=TRUE}
# ANOVA test: A + B vs Control
m2 <- aov(Results ~ Herbicide_groups, 
          data = long_data)
summary.aov(m2, split=list(Herbicide_groups=list("A + B vs. Control."=2)))
```
No significant difference against the control group.

### Hypothesis - Linear Contrasts
- $H_0$: $\mu HerbicideA\ + \mu HerbicideB\ = \mu Control$\
- $H_1$: $\mu HerbicideA\ +  \mu HerbicideB\ \ne \mu Control$\

### Test Statistic
- $F_0 = 0.045$

### p-Value
- $p = 0.8322$

## Conclusion
Keep $H_0$ at the $\alpha$=0.05 ($\alpha_B$ = 0.025) level. There is not sufficient evidence to suggest that the average yield between the Herbicide A and Herbicide B groups are different from the control group.

#### 6. Use the appropriate posthoc test to perform all pairwise comparisons of yield. Test at the $\alpha=0.05$ level.

```{r, echo=TRUE}
TukeyHSD(m1)
```

There is a difference between Herbicide A and B (p = 0.022).

#### 7. Use the appropriate posthoc test to only compare the individual herbicides to the control. Test at the $\alpha=0.05$ level.

```{r, echo=TRUE}
DunnettTest(x = long_data$Results,
            g = long_data$Herbicide_groups,
            control = "Control")
```
No significant difference against the control group.