---
title: "One-Way ANOVA - Week 5, Lecture 1"
author: "Hector Gavilanes"
format: html
self-contained: true
editor: source
---

```{r setup, echo = TRUE, warnings = FALSE, message = FALSE}
install.packages(DescTools)
library(tidyverse)
library(gsheet)
library(formattable)
# perform Dunnett Test
library(DescTools) 
```

**An agricultural researcher investigated two new herbicides and their effects on crop yields of tomatoes, which can be demolished by the [tomato hornworm](https://extension.umn.edu/yard-and-garden-insects/tomato-hornworms) in mere hours. There were three areas for the experiment: an area for herbicide A, an area for herbicide B, and an area for no herbicide at all (control). At the end of the growing season, the yields (in bushels per acre) were measured and are available in the following [Google Sheet](https://docs.google.com/spreadsheets/d/1CFyuuWy-4YmVU3wkYuC8Uf3xW3qkpCnpF468NjBAhbY/edit#gid=0).**

```{r, echo=TRUE}
# clear environment
rm(list = ls())

# extract data
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1CFyuuWy-4YmVU3wkYuC8Uf3xW3qkpCnpF468NjBAhbY/edit#gid=0"))

# transform data
clean_data <- data %>% 
  # delete empty rows
  drop_na()
# pivot to long data
long_data <- clean_data %>% 
  pivot_longer(cols = c("Herbicide1", "Herbicide2", "Control"),
               names_to = "Herbicide_groups",
               values_to = "Results")
```

#### 1. Construct the 95% confidence interval for the average yield for each group (A, B, control).

```{r, echo=TRUE}
# t.test(clean_data$Control)
# t.test(clean_data$Herbicide1)
# t.test(clean_data$Herbicide2)
# f02 // get confidence intervals for each group
get_ci <- function(df, category, continue_variable){
  ci <- df %>% 
    group_by(!! sym(category)) %>%
    summarise(
      ci_lower = t.test(!! sym(continue_variable), conf.level = 0.95)$conf.int[1],
      ci_upper = t.test(!! sym(continue_variable), conf.level = 0.95)$conf.int[2]
    )
  return(ci)
}
```

#### 2. Challenge! Construct a graph to display all confidence intervals.

```{r, echo=TRUE}
boxplot(clean_data)
boxplot(long_data$Results ~ long_data$Herbicide_groups)
# f03 // get mean and confidence intervals for each group
get_mean_ci <- function(df, category, continue_variable){
  mean_ci <- df %>% 
    group_by(!! sym(category)) %>%
    summarise(
      mean_herbicide = mean(!! sym(continue_variable)),
      ci_lower = t.test(!! sym(continue_variable), conf.level = 0.95)$conf.int[1],
      ci_upper = t.test(!! sym(continue_variable), conf.level = 0.95)$conf.int[2]
    ) %>% 
  return(mean_ci)
}


# f04 // graph confidence intervals for each group
plot_mean_ci <- function(mean_ci){
  mean_ci%>% 
    ggplot(aes(x = Herbicide_groups, y = Results, fill = factor(Herbicide_groups)))+
    geom_bar(stat = "identity")+
    stat_summary(y = mean_herbicide)
    scale_fill_manual(values=c("#c2d4dd", "#b1cbbb", "#e4d1d1"))+
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), size = 1.5, color = "#034f84", width = 0.4)+
    geom_text(aes(y = ci_lower, label = round(ci_lower,1)), size = 4, vjust = 2, color = "black")+
    geom_text(aes(y = ci_upper, label = round(ci_upper,1)), size = 4, vjust = -1, color = "black")+
    labs(title = "Chart with 95% Confidence Intervals")+
    labs(x="Herbicide Groups", y="Yield Results")+
    ylim(0, 100)+
    lims(colour = c("1", "2", "Control"))+
    theme_bw()+
    theme(legend.position="none")
} 
# plot_mean_ci(get_mean_ci(long_data, "Herbicide_groups", "Results"))

# boxplot(get_mean_ci(long_data, "Herbicide_groups", "Results"))
```

#### 3. Challenge! Why should we be cautious when using the individual CIs found in Q1 and displayed in Q2 to directly compare the three groups?
Visually, there is not a significant variance between the groups. However, the ANOVA test signals a difference.
#### 4. Use the appropriate hypothesis test to show that there is a difference in the average yield between the three groups. Test at the $\alpha=0.05$ level.

```{r, echo=TRUE}
# one-way ANOVA model
module_data_anova <- aov(long_data$Results ~ long_data$Herbicide_groups)
summary(module_data_anova)
# Alternative code for ANOVA
# module_1 <- lm(long_data$Results ~ long_data$Groups)
# anova(module_1)
```
### Hypothesis
- $H_0$: $\mu Control = \mu Herbicide1\ = \mu Herbicide2$\
- $H_1$: at least one is different

### Test Statistic
- $F_0 = 3.677$

### p-Value
- $p = 0.0293$

## Conclusion
Reject $H_0$ at the $\alpha$=0.05 level. There is sufficient evidence to suggest that there is a difference in the average yield between the three groups.
#### 5. Construct and test the following contrasts for the following comparisons:

##### a. Herbicide A vs. Herbicide B. Test at the $\alpha=0.05$ level.

```{r, echo=TRUE}
# plot(module_data_anova)

long_data$Herbicide_groups <- as.factor(long_data$Herbicide_groups)
contrasts(long_data$Herbicide_groups)

contrasts(long_data$Herbicide_groups) <- cbind(c(0, -1, 1), 
                                    c(-1, 0.5, 0.5))
```

##### b. Herbicide A and Herbicide B vs. Control. Test at the $\alpha=0.05$ level.

```{r, echo=TRUE}
m1 <- aov(Results ~ Herbicide_groups, 
          data = long_data)
summary.aov(m1, split=list(Herbicide_groups=list("A vs. B"=1,
                                                 "A + B vs. Control."=2)))
```

#### 6. Use the appropriate posthoc test to perform all pairwise comparisons of yield. Test at the $\alpha=0.05$ level.

```{r, echo=TRUE}
TukeyHSD(m1)
plot(TukeyHSD(m1), las = 2)

```

There is a difference between Herbicide 1 and 2 (p = 0.022).

#### 7. Use the appropriate posthoc test to only compare the individual herbicides to the control. Test at the $\alpha=0.05$ level.

```{r, echo=TRUE}
DunnettTest(x = long_data$Results,
            g = long_data$Herbicide_groups,
            control = "Control")
```
No herbicide groups are significant different from the control group.