---
title: "Project 3"
author: "Your Name Here"
format: html
self-contained: true
editor: source
---

```{r setup, echo = TRUE, warnings = FALSE, message = FALSE}

```

**Consider the insurance cost data availabile [here](https://github.com/stedy/Machine-Learning-with-R-datasets/blob/master/insurance.csv).**

**1. Import the data. Extra credit will be given to students that figure out how to directly import from GitHub.**

```{r, echo = TRUE}

```

**2. Fill in the following table:**

|                    |        | Mean (Standard Deviation) or Number (Percent) |
|--------------------|--------|-----------------------------------------------|
| Charges            |        |                                               |
| Age                |        |                                               |
| BMI                |        |                                               |
| Sex                | Male   |                                               |
|                    | Female |                                               |
| Smoking Status     | Yes    |                                               |
|                    | No     |                                               |
| Number of Children | 0      |                                               |
|                    | 1      |                                               |
|                    | 2      |                                               |
|                    | ... (add a row for each additional number needed)      |                                               |

```{r, echo = TRUE}

```

**3. Model insurance charges as a function of age, BMI, and number of children. Remember to state the resulting model.**

```{r, echo = TRUE}

```

**4. Provide brief and appropriate interpretations for all regression coefficients.**

**5. Fill in the following table:**

| Predictor          | $\hat{\beta}_i$ (95% CI) | *p*-Value |
|--------------------|--------------------------|-----------|
| Age                |                          |           |
| BMI                |                          |           |
| Number of Children |                          |           |

```{r, echo = TRUE}

```

**(yes, I am asking you to place both the estimated $\beta$ as well as the corresponding 95% CI in the cell.)**

**6. Which, if any, are significant predictors of insurance charges? Test at the $\alpha=0.05$ level. You do not need to state all hypothesis test pieces, but you must provide appropriate justification for your conclusions.**

```{r, echo = TRUE}

```

**7. Use the appropriate hypothesis test to determine if this is a significant regression line. Test at the $\alpha=0.05$ level.**

```{r, echo = TRUE}

```

**8. Construct the correlation matrix for the variables in the regression model. Are any suspiciously high?**

```{r, echo = TRUE}

```

**9. Check for outliers. How many are there?**

```{r, echo = TRUE}

```

**10. Check for influential/leverage points. How many are there?**

```{r, echo = TRUE}

```

**11. Check for multicollinearity. Do the results surprise you?**

```{r, echo = TRUE}

```

**12. Construct a graph to aid with explanation of the regression model. Create lines for 0, 2, and 4 children. You pick what goes on the *x*-axis and what is plugged in for the remaining variable. Extra credit if you make the outlier dots a different color than the non-outlier dots.**

```{r, echo = TRUE}

```

**13. Write a short paragraph to accompany your results, appropriate for your supervisor who is not a statistician or data scientist. Outline your modeling technique as well as the summary of the data (i.e., the first table) and results.**