---
title: "One and Two Proportions - Week 10, Lecture 1"
author: "Hector Gavilanes"
format: html
self-contained: true
editor: source
---
#### Load libraries
```{r setup, echo = TRUE, warnings = FALSE, message = FALSE}
library("tidyverse")
library("gsheet")
library("formattable")
```

**1. Consider the Jackson Heart Study dataset.**

```{r}
# clear environment
rm(list = ls())
# import data
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1H3TP-2SBMGleriJLESOe1cdCjtSj2F76bUh5iBqC8tI/edit#gid=2142000894"))
```

**1a. Find the 95% CI for the proportion of JHS participants that drink alcohol on a regular basis (alc; 1=yes, 0=no).**
```{r}
# alcohol count
formattable(data) %>% 
  count(alc)

# get Confidence Intervals
prop.test(x = 1271,
          n = 2643, #sample size
          correct=FALSE)
```
- The 95% CI for $\pi$, the population proportion of JHS participants that drink alcohol on a regular basis is (0.462, 0.499)

**1b. Use the appropriate hypothesis test to determine if less than half of the JHS participants drink alcohol on a regular basis. Test at the  level.**

```{r}
# n = 2643
# yes = 1271
# no = 1372

# less than 50% drink
prop.test(x = 1271,
          n = 2643+1271, #sample size
          p = 0.50,
          alternative = "less",
          correct=FALSE)

z0 = round(sqrt(480.94), digits = 3)
z0
```
### Hypothesis Test: Alcohol Drinkers

### Hypotheses

- $H_0: \pi_{alcohol} \le 0.50$
- $H_1: \pi_{alcohol} > 0.50$

### Test Statistic and p-Value

- $X^2_{0}$ = 480.94 or $z_0$ = $\sqrt{3.859}$ = `r z0` 

### p-Value

- $p < 0.001$

## Conclusion

Reject $H_0$ at $\alpha = 0.05$. There is sufficient evidence to suggest that less than 50% of the JHS participants drink alcohol on a regular basis.

**1c. Find the 99% CI for the difference in the proportion of male and female JHS participants that drink alcohol.**
```{r}
# get matrix/table
table(data$alc, data$sex)

# n female = 1666
982+684

# n male = 977
390+587

# 99% CI for two proportions
prop.test(x = c(684, 587),
          n = c(1666, 977),
          correct = FALSE,
          conf.level = 0.99)
```
- The 99% CI for $\pi_F$ - $\pi_M$ is (-0.241, -0.139) 

1d. Use the appropriate hypothesis test to determine if there are fewer drinkers with hypertension (HTN; 1=yes, 0=no) than those without hypertension. Test at the  level.
```{r}
# count hypertension
data %>% 
  count(data$alc)
# drinkers with hypertension count
table(data$HTN, data$alc)

# less drinkers with hypertension
# drinkers n = 1271
# drinkers+HTN = 588
# drinkers-HTN = 683
prop.test(x = 588,
          n = 1271+588,
          alternative = "less",
          correct = FALSE)

# get Z0
z_htn = round(sqrt(250.94), digits = 3)
z_htn
```
### Hypothesis Test: Alcohol Drinkers

### Hypotheses

- $H_0: \pi_{hypertension} \ge \pi_{alcohol}$
- $H_1: \pi_{hypertension} < \pi_{alcohol}$

### Test Statistic and p-Value

- $X^2_{0}$ = 250.94 or $z_0$ = $\sqrt{250.94}$ = `r z_htn` 

### p-Value

- $p < 0.001$

## Conclusion

Reject $H_0$ at $\alpha = 0.05$. There is sufficient evidence to suggest that there are fewer drinkers with hypertension than those without hypertension.

1e. Challenge! Construct a graph to help with explanation to the lead scientist of JHS.
```{r}
# data types
class(data$sex)
class(data$alc)
class(data$HTN)
# factor
sex_factor <- as.factor(data$sex)
alc_factor <- as.factor(data$alc)
htn_factor <- as.factor(data$HTN)

pie(table(data$alc))
pie(table(data$sex))
pie(table(data$HTN))
```

**2. A large retail lawn care dealer currently provides a 2-year warranty on all lawn mowers sold at its stores. A new employee suggested that the dealer could save money by just not offering the warranty. To evaluate this suggestion, the dealer randomly decides whether or not to offer the warranty to the next 50 customers who enter the store and express an interest in purchasing a lawnmower. Out of the 25 customers offered the warranty, 10 purchased a mower as compared to 4 of 25 not offered the warranty.

2a. Find a 95% CI for , the difference in the proportions of customers purchasing lawnmowers with and without the warranty.

2b. What are the requirements for using the z distribution when analyzing proportions?

2c. Check those requirements here.

2d. Is the CI in part a valid? Why or why not?