---
title: "Two-Sample Means (Dependent) - Week 2, Lecture 2"
author: "Hector Gavilanes"
format: html
self-contained: true
editor: source
---

```{r setup, echo = TRUE, warnings = FALSE, message = FALSE}
library(tidyverse)
library(gsheet)
```

**A new 2 hour reading comprehension program has been developed by a local tutoring center. A pilot study of 30 students has been conducted and we have data on a reading comprehension test taken before the program and a separate reading comprehension test taken after the program. Consider the data, found in the following Google Sheet.**

```{r, echo = TRUE}
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1ijxX7c6qcFAw-r_mNSxC-Ps5b5njw-WtE4RcxPTySXU/edit?usp=sharing"))
```

#### 1. Summarize the data. Provide appropriate summary statistics for the study population.

```{r, echo=TRUE}
summary(data)
```

#### 2. Construct the 95% confidence interval for the difference in test scores. Hint: we want to discuss this in terms of the *increase* in test scores.

```{r, echo=TRUE}
t.test(data$Pre, data$Post,
       data = data,
       paired = TRUE,
       conf.level = 0.95)
```

The 95% CI is (-5.37, -2.23)

### Hypothesis
H0: $\mu_Pre = \mu_Post$\
H1: $\mu_Pre \ne \mu_Post$\

### Test
t0 = -4.95

### p-Value
p = 2.918e-05

## Conclusion
Reject H0 at $\alpha$=.05 level.There is a difference in the test scores. 

#### 3. Use the appropriate hypothesis test to determine if the program raises comprehension test averages by at least 5 points. Test at the &alpha;=0.05 level.

```{r, echo=TRUE}
t.test(data$Pre, data$Post,
       data = data,
       paired = TRUE,
       alternative = "greater",
       conf.level = 0.95)
```
### Hypothesis
H0: $\mu_Pre \le \mu_Post$\
H1: $\mu_Pre > \mu_Post$\

### Test
t0 = -11.46

### p-Value
p = 1

## Conclusion
Keep H0 at the $\alpha$=0.05 level. There is sufficient evidence to suggest that the program raises comprehension test averages by at least 5 points.

#### 4. (Challenge!) Construct a graph that will help you explain your findings to the tutoring center.

```{r, echo=TRUE}
qq_graph <- data %>% 
  mutate(diff = Pre - Post) %>% 
  ggplot(aes(sample = diff))+
  stat_qq(size = 3)+
  stat_qq_line()+
  theme_bw()+
  xlab("Theoretical")+
  ylab("Sample")
qq_graph

```

#### 5. Write a brief paragraph explaining your findings to the tutoring center. 









