---
title: "Project 2"
author: "Hector Gavilanes"
format: html
self-contained: true
editor: source
---

```{r setup, echo = TRUE, warnings = FALSE, message = FALSE}
library(tidyverse)
library(gsheet)
library(formattable)
library(DescTools) # perform Dunnett, Levene Test
library(pgirmess) # perform Kruskal-Wallis Post-hoc
library(ggpubr) # interaction plot
library(ggstatsplot) # boxplot
library(ISLR)
```

**1. Consider a sample of data from the [Jackson Heart Study](https://www.jacksonheartstudy.org/). In this problem, we will be examining body mass index (*BMI*; kg/m<sup>2</sup>) as a function of health as categorized by physical activity (*PA3cat*; Ideal Health, Intermediate Health, and Poor Health).**

```{r}
# clear environment
rm(list = ls())
# import data
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1H3TP-2SBMGleriJLESOe1cdCjtSj2F76bUh5iBqC8tI/edit#gid=2142000894"))
# clean data
clean_data <- data %>% 
  select(BMI, BMI3cat) %>% 
  drop_na()

# ---CUSTOM COLORS---
custom_green_dark = "#71CA97"
custom_green_light = "#DeF7E9"
custom_gray = "#3D3D3D"
custom_orange = "#FC7307"
custom_purple = "#5A18C9"
# custom_tan_light = "#F7E3D4"
custom_red_light = "#F7D0C7"
custom_red_dark = "#F1917C"
# custom_density_hue = c(custom_purple,custom_orange)
# ---FUNCTIONS---
# f00 - Change default table format
change_table_format <- function(data_table){
  # local variables
  first_column <- data_table[1]
  dataframe_length <- NCOL(data_table)
  # format
  formattable(data_table, align = c("l",rep("c", dataframe_length - 1)),
              list(first_column = formatter("span", style = ~ style(color = custom_gray, font.weight = "bold")),
              area(row = 1, col = 2:dataframe_length) ~ color_tile(custom_green_light, custom_green_dark),
              area(row = 3, col = 2:dataframe_length) ~ color_tile(custom_red_light, custom_red_dark)
              )#/end list
              )#/end formattable
}
# f01 - get custom summary statistics
# !! sym() creates a symbol from a string
get_custom_summary <- function(data_table, category, values){
  summary <- data_table %>%
    group_by(!! sym(category)) %>% 
    summarise(Avg = mean(!! sym(values)),
              Median = median(!! sym(values)),
              SD = sd(!! sym(values)),
              Min = min(!! sym(values)),
              Max = max(!! sym(values)),
              IQRange = IQR(!! sym(values))
              )
    return(summary)
}
```


**a. Find the appropriate summary statistics to summarize the data.**
```{r}
change_table_format(get_custom_summary(clean_data, "BMI3cat", "BMI"))
```

**b. Use ANOVA to determine if there is a difference in BMI between the three levels of health as categorized by physical activity. Test at the $\alpha=0.05$ level.**
```{r}
# One-way ANOVA
# Null Hypothesis:
# - H0: All means of observations grouped by BMI category are the same
# - H1: At least one category is different

h1 <- aov(clean_data$BMI ~ clean_data$BMI3cat)
summary(h1)

```
### Hypothesis
- $H_0$: $\mu Ideal = \mu Intermediate\ = \mu Poor$\
- $H_1$: at least one is different

### Test Statistic
- $F_0 = 1966$

### p-Value
- $p < 0.001$

## Conclusion
Reject $H_0$ at the $\alpha$=0.05 level. There is sufficient evidence to suggest that there is a difference in BMI between the three levels of health.

**c. Use the Kruskal-Wallis to determine if there is a difference in BMI between the three levels of health as categorized by physical activity (*PA3cat*; Ideal Health, Intermediate Health, and Poor Health). Test at the $\alpha=0.05$ level.**
```{r}
kruskal.test(clean_data$BMI ~ clean_data$BMI3cat)
```
### Hypothesis: Kruskal-Wallis
- $H_0$: $M_{Ideal} = M_{Intermediate} = M_{Poor}$\
- $H_1$: at least one group is different

### Test Statistic
- $H = 2129.4$

### p-Value
- $p < 0.001$

## Conclusion
Reject $H_0$ at the $\alpha$=0.05 level. There is sufficient evidence to suggest that there is a difference in BMI between the three levels of health.

**d. State/explain the ANOVA assumptions and assess them. Remember to state your conclusion (that you either meet the ANOVA assumptions or you do not).**

We are unwilling to assume normality because we have found the assumption of normality invalid in diagnostic checks. We have performed the following ANOVA assumptions:

1. Independence – the observations in each group need to be independent of each other.
The data provided meets this requirement.

2. Normality – the dependent variable should be approximately normally distributed for each combination of the groups.
However, the residuals are not normally distributed; thus, we cannot assume normality.

```{r}
# 2. Normality check
plot(h1)
boxplot(clean_data$BMI ~ clean_data$BMI3cat)
```
3. Equal Variance - the variances for each group are equal or approximately equal.
Since the p-value < 0.001 which is lower than our 0.05 level of significance, we cannot assume equality of variances.

```{r}
# 3. Levene's test for equality of variance
LeveneTest(clean_data$BMI ~ clean_data$BMI3cat,
           exact = FALSE)
```



**e. Based on your responses in part (d), which test's result are you going to present to the lead scientist at the JHS? (Hint: you will pick either (b) or (c)).**

Nonparametric test, Kruskal-Wallis, in this case.

**f. Perform the appropriate posthoc test to determine pairwise differences between the three groups. Test at the $\alpha=0.05$ level.**
```{r}
# Kruskal-Wallis post-hoc
kruskalmc(clean_data$BMI ~ clean_data$BMI3cat)

# Multiple pairwise-comparison between groups
pairwise.wilcox.test(clean_data$BMI, clean_data$BMI3cat,
                     p.adjust.method = "bonferroni",
                     exact = FALSE)
# nonparametric DunnTest
DunnTest(clean_data$BMI ~ clean_data$BMI3cat)
```

**g. Write a brief paragraph describing the results of your analysis. Include summary statistics, results of your chosen hypothesis test from part (e), results of the posthoc test in part (f), and a graph to help convey the results.**

## Analisis
The ANOVA assumptions were invalid after normality checks and Levene’s test.
Therefore, to find a statistically significant difference in BMI between the three levels of health (poor, intermediate, and ideal), a nonparametric Kruskal-Wallis test was conducted.
The procedure following a Kruskal-Wallis test was to perform a Dunn's posthoc test.

## Conclusion
All three levels of health grouped by BMI were significantly different from each other. BMI and physical activity displayed an inverse correlation. People with the lowest BMI had a higher physical activity, categorized as ideal health. On the other hand, people with the highest BMI calculations had lower physical activity. 

54% of the sample represented the dominant group with poor health levels of physical activity.


### Summary Statistics

To better understood the data, we calculated the mean, median, and standard deviation.
```{r, echo=FALSE}
change_table_format(get_custom_summary(clean_data, "BMI3cat", "BMI"))
```

### ANOVA Assumptions

We were unwilling to assume normality because we have found the assumption of normality invalid in diagnostic checks. We have performed the following ANOVA assumptions:

* **1. Independence** – the observations in each group need to be independent of each other.
The data provided met this requirement.

* **2. Normality** – the dependent variable should be approximately normally distributed for each combination of the groups.
However, the residuals are not normally distributed; thus, normality was not assumed.

The *Fitted vs. Residuals* graph showed how the residuals had differences in size among the groups.

```{r, echo=FALSE}
# 2. Normality check
lm_model = lm(clean_data$BMI ~ clean_data$BMI3cat)
#anova(lm_model)

# fitted vs residuals
ggplot(lm_model) + 
  geom_point(aes(x=.fitted, y=.resid), 
             size=3, alpha = 0.35, color = custom_purple)+
  labs(title = "Fitted vs Residuals")+
  xlab("Fitted")+
  ylab("Standarized Residuals")+
  theme_bw()
```

The *QQ (quantile-quantile) plot*  did not display the normality of residuals. The residuals did not follow the 45-degree line.
```{r, echo=FALSE}
# QQ plot
ggplot(lm_model, aes(sample = lm_model$residuals))+
  geom_qq(size=3, alpha = 0.6, color = custom_orange)+
  geom_qq_line()+
  theme_bw() +
  labs(title = "Normal Q-Q Plot")+
  xlab("Theoretical Quantiles") +
  ylab("Standarized Residuals")
```

Moreover, a *Boxplot* displayed the distribution of BMI health for each group with significant differences.
```{r, echo=FALSE}
ggbetweenstats(
  data = clean_data,
  x = BMI3cat,
  y = BMI,
  plot.type = "box",
  type = "np",
  pairwise.display = "s",
  var.equal = T,
  p.adjust.method = "bonferroni",
  ylab = "BMI",
  xlab = "Health Levels",
  k = 0,
  bf.message = F
)
```
* **3. Equal Variance** - the variances for each group are equal or approximately equal.

Since the p-value < 0.001 was lower than 0.05 level of significance, we did not assume equality of variances.

```{r, echo=FALSE, warning=FALSE}
# 3. Levene's test for equality of variance
LeveneTest(clean_data$BMI ~ clean_data$BMI3cat,
           exact = FALSE)
```
### Kruskal-Wallis Test
We used a nonparametric alternative, Kruskal-Wallis, to ANOVA test. 
```{r, echo=FALSE}
kruskal.test(clean_data$BMI ~ clean_data$BMI3cat)
```
#### Hypothesis: 
- $H_0$: $M_{Ideal} = M_{Intermediate} = M_{Poor}$\
- $H_1$: at least one group is different

#### Test Statistic
- $H = 2129.4$

#### p-Value
- $p < 0.001$

#### Conclusion
Reject $H_0$ at the $\alpha$=0.05 level. There was sufficient evidence to suggest that there was a difference in BMI between the three levels of health.

### Posthoc Dunn's Test
Dunn's test determined which groups were significantly different with comparisons between each independent group.

All groups had significant variance among each other, with a p-value < 0.001 lower than 0.05 level of significance.
```{r, echo=FALSE}
# nonparametric DunnTest
DunnTest(clean_data$BMI ~ clean_data$BMI3cat)
```


**2. Again, consider the sample of data from the JHS. In this problem, we will be examining systolic blood pressure (*sbp*; mmHg) as a function of health as categorized by body mass index (*BMI3cat*; Ideal Health, Intermediate Health, and Poor Health), smoking status (*everSmoker*; 0=never smoker, 1=former or current smoker), and the interaction between health as categorized by body mass index and smoking status.**

```{r}
blood_data <- data %>% 
  select(BMI3cat, sbp, everSmoker) %>% 
  drop_na()
```

**a. Find the appropriate summary statistics to summarize the data.**
```{r}
# change_table_format(get_custom_summary(blood_data, "everSmoker", "BMI"))
bmi_smoke_mean <- tapply(blood_data$sbp, INDEX = list(blood_data$BMI3cat, blood_data$everSmoker), FUN = mean)
formattable(bmi_smoke_mean)
# attributes(bmi_smoke_mean)
```

**b. Use ANOVA to determine if there is a difference in systolic blood pressure between health as categorized by body mass index, smoking status, and the interaction between health as categorized by body mass index and smoking status. Test at the $\alpha=0.05$ level. Remember to remove the interaction term if appropriate to do so.**
```{r}
# Two-way ANOVA with Interaction
# Null Hypothesis:
# - H1: The means of observations grouped by BMI are the same
# - H2: The means of observations grouped by Smoking are the same
# - H3: There is no interaction between BMI and Smoking

summary(h3_interaction <- aov(blood_data$sbp ~ blood_data$BMI3cat * factor(blood_data$everSmoker)))
```
Reject H1.
Reject H2.
Reject H3.

No further test is required because there is an interaction.
After removing the interaction term, the ANOVA table was reconstructed.
```{r}
# Two-way ANOVA
summary(two_way <- aov(blood_data$sbp ~ factor(blood_data$BMI3cat) + factor(blood_data$everSmoker)))
```

**c. Construct a profile plot to help explain results of part (b).**

```{r}
# profile plot using ggline
profile <- ggline(blood_data, x = "BMI3cat", y = "sbp", color = "everSmoker",
       add = "mean",
       show.line.label = FALSE,
       palette = c("magenta", "gray"),
       xlab = "BMI Health Categories",)+
  theme_bw()
  
interaction_plot <- ggpar(profile, 
                      main = "Interaction Plot: \n Systolic Blood Pressure & Smoking Status",
                      legend.title = "Smoking Status",
                      legend = "right",
                      ylim = c(20, 40))
interaction_plot

# smoke
# profile plot using ggline
profile <- ggline(blood_data, x = "everSmoker", y = "sbp", color = "BMI3cat",
       add = "mean",
       show.line.label = FALSE,
       palette = c("magenta", "gray", "orange"),
       xlab = "BMI Health Categories",)+
  theme_bw()
```

**d. Use the appropriate posthoc test to determine differences in main effects or the interaction term, whichever is appropriate for the results in part (b). Test at the $\alpha=0.05$ level.**

## Levene's Test for Equality of Variances
```{r}
LeveneTest(blood_data$BMI ~ blood_data$BMI3cat * blood_data$everSmoker)
```
## ANOVA Assumption
```{r}
plot(h3_interaction)
```


## Main Effects for BMI Health
```{r}
kruskal.test(blood_data$BMI ~ blood_data$BMI3cat)
kruskalmc(blood_data$BMI ~ blood_data$BMI3cat)
pairwise.wilcox.test(blood_data$BMI, blood_data$BMI3cat,
                     p.adjust.method = "bonferroni",
                     exact = FALSE)
DunnTest(blood_data$BMI ~ blood_data$BMI3cat)

```

## Main Effects for Smoking Health
```{r}
kruskal.test(blood_data$BMI ~ blood_data$everSmoker)
kruskalmc(blood_data$BMI ~ blood_data$everSmoker)
pairwise.wilcox.test(blood_data$BMI, blood_data$everSmoker,
                     p.adjust.method = "bonferroni",
                     exact = FALSE)
DunnTest(blood_data$BMI ~ blood_data$everSmoker)
```

**e. Write a brief paragraph describing the results of your analysis. Include summary statistics, results of the hypothesis testing in part (b), results of the posthoc test in part (d), and the profile plot in part (c) to help convey the results.**

