---
title: "Multiple Regression: Diagnostics - Week 10, Lecture 1"
author: "Hector Gavilanes"
format: html
self-contained: true
editor: source
---

**Recall the airline data from last week; we modeled revenue as a function of distance and population. The data are given here: [Google Sheet](https://docs.google.com/spreadsheets/d/10ZN0jMYdVsG2ucIJFSTquVRW916Jj-JmZIGQuqSQhy0/edit?usp=sharing).**

## Libraries
```{r setup, echo = TRUE, warnings = FALSE, message = FALSE}
library(tidyverse)
library(gsheet)
library(lindia) # Influence & Leverage: gg_cooksd()
library(car) #multicollinearity: vif()
library(olsrr) #multicollinearity: vif()
library(ggpubr) #side-by-side graphs
```

## Data
```{r}
# clear environment
rm(list = ls())

# extract data
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/10ZN0jMYdVsG2ucIJFSTquVRW916Jj-JmZIGQuqSQhy0/edit?usp=sharing"))
```

1. Summarize your findings from last week. What was the resulting model, which were significant predictors of revenue (no formal hypothesis test needed - just state with appropriate evidence)? Also include the graph constructed to visualize the model.

#### Summary
```{r}
# Summary
data %>% 
  select(Revenue, Distance, Population) %>% 
  summary(data)
data %>% 
  select(Population) %>% 
  summarise(sd = sd(Population))
```


#### Correlation
```{r}
# calculate Pearson's correlation matrix
cor(data,
    method = "pearson",
    use = "complete.obs")
```

There is a correlation between Revenue and Population which is 0.762.

```{r}
airport_model <- lm(data$Revenue ~ data$Population + data$Distance)
summary(airport_model)
```

## Model formula:
$$\hat{y} = 144.39 + 1.09{\text{ population}} + 0.17{\text{ distance}} $$
- Population is a significant predictor with $p < 0.001$.
- Distance is not a significant predictor with $p = 0.279$.
- The regression model is significant with $p < 0.001$

#### Graph the Multiple Linear Model

#### Visualizing Population
```{r}
# get coefficients
c1 <- coefficients(airport_model)
c1

# Population on the X-axis
quantile(data$Distance, na.rm = TRUE)

# create lines graph for Population of 150, 211, 300
data <- data %>%
  mutate(p150 = c1[1] + c1[2]*Population + c1[3]*150,
         p211 = c1[1] + c1[2]*Population + c1[3]*211,
         p300 = c1[1] + c1[2]*Population + c1[3]*300)

# Population ggplot 
gg_population <- data %>% ggplot(aes(x = Population, y = Revenue)) +
  geom_point(size=3, color = "darkgray") + 
  geom_line(aes(y = p150), color = "orange") +
  geom_line(aes(y = p211), color = "magenta") +
  geom_line(aes(y = p300), color = "blue") +
  geom_text(aes(x =264, y = 460, label = "30,000 persons")) +
  geom_text(aes(x = 264, y = 445, label = "21,100 persons")) +
  geom_text(aes(x = 264, y = 430, label = "15,000 persons")) +
  labs(title = "Multiple Linear Regression",
       subtitle = "Predictor: Population",
       x = "Population",
       y = "Revenue") +
  xlim(c(50, 275))+
  ylim(c(200, 480))+
  theme_bw()
gg_population
```
#### Visualizing Distance
```{r}
# Distance on the X-axis
quantile(data$Population, na.rm = TRUE)

# create lines graph for Distance of 55, 74, 150
data <- data %>%
  mutate(d55 = c1[1] + c1[2]*55 + c1[3]*Distance,
         d74 = c1[1] + c1[2]*74 + c1[3]*Distance,
         d150 = c1[1] + c1[2]*150 + c1[3]*Distance)

# Distance ggplot 
gg_distance <- data %>% ggplot(aes(x = Distance, y = Revenue)) +
  geom_point(size=3, color = "darkgray") + 
  geom_line(aes(y = d55), color = "orange") +
  geom_line(aes(y = d74), color = "magenta") +
  geom_line(aes(y = d150), color = "blue") +
  geom_text(aes(x = 354, y = 370, label = "150 miles")) +
  geom_text(aes(x = 354, y = 287, label = "74 miles")) +
  geom_text(aes(x = 354, y = 265, label = "55 miles")) +
  labs(title = "Multiple Linear Regression",
       subtitle = "Predictor: Distance",
       x = "Distance",
       y = "Revenue") +
  xlim(c(120, 365))+
  ylim(c(200, 450))+
  theme_bw()
gg_distance
```


2. Determine if there are outliers in this model. Be sure to count and list the observations that are outliers.
```{r}
n = 2.5
data <- data %>% 
  mutate(Outlier = abs(rstandard(airport_model)) > n)
data %>% count(Outlier)

outlier_df <- data %>% 
  select(Revenue, Distance, Population, Outlier) %>% 
  drop_na() %>% 
  mutate(obs = row_number()) 
outlier_df %>% 
  filter(Outlier == TRUE)
```

3. Determine if there are any influence and leverage points in this model. Be sure to count and list the observations that are considered influential and leverage.
```{r}
gg_cooksd(airport_model) + theme_bw()
ols_plot_resid_lev(airport_model)
```
- Points 20, 21, 22 need further investigation.

4. Is there evidence of multicollinearity? 
```{r}
# vif(airport_model)
ols_vif_tol(airport_model)
```
- No multicollinearity is present.

5. Perform a sensitivity analysis: remove the data points identified as potential offenders in Q2 and Q3 and reconstruct the model. Be sure to state the resulting model.
```{r}
sensitivity_df <- outlier_df %>% 
  filter(Outlier == FALSE & !obs %in% c(20, 21, 22))
# data$Revenue ~ data$Population + data$Distance
sensitivity_model <- lm(sensitivity_df$Revenue ~ sensitivity_df$Population + sensitivity_df$Distance)
summary(sensitivity_model)
```
## Sensitivity Model formula:
$$\hat{y} = 128.89 + 0.83{\text{ population}} + 0.35{\text{ distance}} $$
6. Using the model in Q5, are there any significant predictors of revenue?

- Population is a significant predictor with $p = 0.0310$.
- Distance is a significant predictor with $p = 0.0347$.
- The regression model is significant with $p = 0.0343$

7. Reconstruct the data visualization from last week, but using the new model's results.
#### Graph the Sensitivity Multiple Linear Model

#### Visualizing Distance
```{r}
# get coefficients
c1 <- coefficients(sensitivity_model)

# Distance on the X-axis
quantile(sensitivity_df$Population, na.rm = TRUE)

# create lines graph for Distance of 55, 73, 95
sensitivity_df <- sensitivity_df %>%
  mutate(d55 = c1[1] + c1[2]*55 + c1[3]*Distance,
         d73 = c1[1] + c1[2]*73 + c1[3]*Distance,
         d95 = c1[1] + c1[2]*95 + c1[3]*Distance)

# Distance sensitivity ggplot 
gg_sensitivity_distance <- sensitivity_df %>% ggplot(aes(x = Distance, y = Revenue)) +
  geom_point(size=3, color = "darkgray") + 
  geom_line(aes(y = d55), color = "orange") +
  geom_line(aes(y = d73), color = "magenta") +
  geom_line(aes(y = d95), color = "blue") +
  geom_text(aes(x = 253, y = 295, label = "95 miles")) +
  geom_text(aes(x = 253, y = 275, label = "73 miles")) +
  geom_text(aes(x = 253, y = 260, label = "55 miles")) +
  labs(title = "Sensitivity Analysis",
       subtitle = "Predictor: Distance",
       x = "Distance",
       y = "Revenue") +
  theme_bw()
  
gg_sensitivity_distance

distance_comparison <- gg_sensitivity_distance +
  xlim(c(120, 365))+
  ylim(c(200, 450))
```

#### Visualizing Population
```{r}
# Population on the X-axis
quantile(sensitivity_df$Distance, na.rm = TRUE)

# create lines graph for Population of 150, 209, 240
sensitivity_df <- sensitivity_df %>%
  mutate(p150 = c1[1] + c1[2]*Population + c1[3]*150,
         p209 = c1[1] + c1[2]*Population + c1[3]*209,
         p240 = c1[1] + c1[2]*Population + c1[3]*240)

# Sensitivity Population ggplot
gg_sensitivity_population <- sensitivity_df %>% ggplot(aes(x = Population, y = Revenue)) +
  geom_point(size=3, color = "darkgray") + 
  geom_line(aes(y = p150), color = "orange") +
  geom_line(aes(y = p209), color = "magenta") +
  geom_line(aes(y = p240), color = "blue") +
  geom_text(aes(x =106, y = 298, label = "24,000 persons")) +
  geom_text(aes(x = 106, y = 285, label = "20,900 persons")) +
  geom_text(aes(x = 106, y = 265, label = "15,000 persons")) +
  labs(title = "Sensitivity Analysis",
       subtitle = "Predictor: Population",
       x = "Population",
       y = "Revenue") +
  theme_bw()

gg_sensitivity_population

population_comparison <- gg_sensitivity_population + 
  xlim(c(50, 275))+
  ylim(c(200, 480))
```

8. Put the graphs from last week and Q7 side by side (hint: use `ggpubr`).

#### Population Comparison
```{r}
ggarrange(distance_comparison, gg_distance, ncol = 2, nrow = 1)
ggarrange(population_comparison, gg_population, ncol = 2, nrow = 1)
```


9. Compare and contrast the modeling results and the graphs that are side by side in Q8.

## Sensitivity Analysis
$$\hat{y} = 144.39 + 1.09{\text{ population}} + 0.17{\text{ distance}} $$
- Population is a significant predictor with $p < 0.001$.
- Distance is not a significant predictor with $p = 0.279$.
- The regression model is significant with $p < 0.001$

$$\hat{y} = 128.89 + 0.83{\text{ population}} + 0.35{\text{ distance}} $$

- Population is a significant predictor with $p = 0.0310$.
- Distance is a significant predictor with $p = 0.0347$.
- The regression model is significant with $p = 0.0343$

10. Why do you think I assigned this particular dataset for analysis? What are you taking away from this assignment?

In this particular assignment, the data has outliers with high influence and high leverage. After removing three outliers, our model is different from the original one.
Further investigation is needed to determine possible problems.

Some true outliers represent natural variations, and they should be left intact in the dataset. Other outliers are problematic, and should be removed because they represent measurement errors, data entry or processing errors, or poor sampling.

