---
title: "Multiple Regression: Diagnostics - Week 10, Lecture 1"
author: "Hector Gavilanes"
format: html
self-contained: true
editor: source
---

**Recall the airline data from last week; we modeled revenue as a function of distance and population. The data are given here: [Google Sheet](https://docs.google.com/spreadsheets/d/10ZN0jMYdVsG2ucIJFSTquVRW916Jj-JmZIGQuqSQhy0/edit?usp=sharing).**

## Libraries
```{r setup, echo = TRUE, warnings = FALSE, message = FALSE}
library(tidyverse)
library(gsheet)
library(lindia) # Influence & Leverage: gg_cooksd()
library(car) #multicollinearity: vif()
library(olsrr) #multicollinearity: vif()
```

## Data
```{r}
# clear environment
rm(list = ls())

# extract data
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/10ZN0jMYdVsG2ucIJFSTquVRW916Jj-JmZIGQuqSQhy0/edit?usp=sharing"))
```

1. Summarize your findings from last week. What was the resulting model, which were significant predictors of revenue (no formal hypothesis test needed - just state with appropriate evidence)? Also include the graph constructed to visualize the model.
```{r}
airport_model <- lm(data$Revenue ~ data$Population + data$Distance)
```

## Model formula:
$$\hat{y} = 144.39 + 1.09{\text{ population}} + 0.17{\text{ distance}} $$
#### Significant Predictor: Population
After adjusted for revenue, there was sufficient evidence to suggest that population was a significant predictor with a $p < 0.001$.

#### Graph the Multiple Linear Model
```{r}
# get coefficients
c1 <- coefficients(airport_model)
c1
# Distance on the X-axis
quantile(data$Population, na.rm = TRUE)
# create lines graph for Distance of 55, 74, 150
data <- data %>%
  mutate(d55 = c1[1] + c1[2]*55 + c1[3]*Distance,
         d74 = c1[1] + c1[2]*74 + c1[3]*Distance,
         d150 = c1[1] + c1[2]*150 + c1[3]*Distance)
```

#### Visualizing Population
```{r}
# Population on the X-axis
quantile(data$Distance, na.rm = TRUE)
# create lines graph for Population of 150, 211, 300
data <- data %>%
  mutate(p150 = c1[1] + c1[2]*Population + c1[3]*150,
         p211 = c1[1] + c1[2]*Population + c1[3]*211,
         p300 = c1[1] + c1[2]*Population + c1[3]*300)
# ggplot graph
data %>% ggplot(aes(x = Population, y = Revenue)) +
  geom_point(size=3, color = "darkgray") + 
  geom_line(aes(y = p150), color = "orange") +
  geom_line(aes(y = p211), color = "magenta") +
  geom_line(aes(y = p300), color = "blue") +
  geom_text(aes(x =260, y = 460, label = "15,000 persons")) +
  geom_text(aes(x = 262, y = 445, label = "7,400 persons")) +
  geom_text(aes(x = 262, y = 430, label = "5,500 persons")) +
  labs(title = "Multiple Linear Regression",
       subtitle = "Predictor: Population",
       x = "Population",
       y = "Revenue") +
  xlim(c(50, 275))+
  theme_bw()
```
#### Visualizing Distance
```{r}
# ggplot graph
data %>% ggplot(aes(x = Distance, y = Revenue)) +
  geom_point(size=3, color = "darkgray") + 
  geom_line(aes(y = d55), color = "orange") +
  geom_line(aes(y = d74), color = "magenta") +
  geom_line(aes(y = d150), color = "blue") +
  geom_text(aes(x = 352, y = 370, label = "150 miles")) +
  geom_text(aes(x = 352, y = 290, label = "74 miles")) +
  geom_text(aes(x = 352, y = 270, label = "55 miles")) +
  labs(title = "Multiple Linear Regression",
       subtitle = "Predictor: Distance",
       x = "Distance",
       y = "Revenue") +
  xlim(c(120, 365))+
  theme_bw()
```


2. Determine if there are outliers in this model. Be sure to count and list the observations that are outliers.
```{r}
n = 2.5
data <- data %>% 
  mutate(Outlier = abs(rstandard(airport_model)) > n)
data %>% count(Outlier)

outlier_df <- data %>% 
  select(Revenue, Distance, Population, Outlier) %>% 
  drop_na() %>% 
  mutate(obs = row_number()) 
outlier_df %>% 
  filter(Outlier == TRUE)
```

3. Determine if there are any influence and leverage points in this model. Be sure to count and list the observations that are considered influential and leverage.
```{r}
gg_cooksd(airport_model) + theme_bw()
```
- Points 20, 21, 22 need further investigation.

4. Is there evidence of multicollinearity? 
```{r}
vif(airport_model)
ols_vif_tol(airport_model)

# collinearity_model <- lm(data$Revenue ~ data$Population * data$Distance)
# ols_vif_tol(collinearity_model)
```
- No multicollinearity is present.

5. Perform a sensitivity analysis: remove the data points identified as potential offenders in Q2 and Q3 and reconstruct the model. Be sure to state the resulting model.
```{r}
sensitivity_df <- outlier_df %>% 
  filter(Outlier == FALSE & !obs %in% c(20, 21, 22))
# data$Revenue ~ data$Population + data$Distance
sensitivity_model <- lm(sensitivity_df$Revenue ~ sensitivity_df$Population + sensitivity_df$Distance)
summary(sensitivity_model)
```
## Sensitivity Model formula:
$$\hat{y} = 128.89 + 0.83{\text{ population}} + 0.35{\text{ distance}} $$

6. Using the model in Q5, are there any significant predictors of revenue?

- Population remains a significant predictor with $p = 0.0310$.
- Distance becomes a significant predictor with $p = 0.0347$.
- The regression model is significant with $p = 0.0343$

7. Reconstruct the data visualization from last week, but using the new model's results.
#### Graph the Multiple Linear Model

#### Visualizing Distance
```{r}
# get coefficients
c1 <- coefficients(sensitivity_model)
# Distance on the X-axis
quantile(sensitivity_df$Population, na.rm = TRUE)
# create lines graph for Distance of 55, 74, 150
sensitivity_df <- sensitivity_df %>%
  mutate(d55 = c1[1] + c1[2]*55 + c1[3]*Distance,
         d73 = c1[1] + c1[2]*73 + c1[3]*Distance,
         d95 = c1[1] + c1[2]*95 + c1[3]*Distance)
# ggplot graph
sensitivity_df %>% ggplot(aes(x = Distance, y = Revenue)) +
  geom_point(size=3, color = "darkgray") + 
  geom_line(aes(y = d55), color = "orange") +
  geom_line(aes(y = d73), color = "magenta") +
  geom_line(aes(y = d95), color = "blue") +
  # geom_text(aes(x = 352, y = 370, label = "150 miles")) +
  # geom_text(aes(x = 352, y = 290, label = "74 miles")) +
  # geom_text(aes(x = 352, y = 270, label = "55 miles")) +
  labs(title = "Multiple Linear Regression",
       subtitle = "Predictor: Distance",
       x = "Distance",
       y = "Revenue") +
  # xlim(c(120, 365))+
  theme_bw()
```
```{r}
# Population on the X-axis
quantile(sensitivity_df$Distance, na.rm = TRUE)
# create lines graph for Population of 150, 211, 300
sensitivity_df <- sensitivity_df %>%
  mutate(p150 = c1[1] + c1[2]*Population + c1[3]*150,
         p211 = c1[1] + c1[2]*Population + c1[3]*211,
         p300 = c1[1] + c1[2]*Population + c1[3]*300)
# ggplot graph
sensitivity_df %>% ggplot(aes(x = Population, y = Revenue)) +
  geom_point(size=3, color = "darkgray") + 
  geom_line(aes(y = p150), color = "orange") +
  geom_line(aes(y = p211), color = "magenta") +
  geom_line(aes(y = p300), color = "blue") +
  # geom_text(aes(x =260, y = 460, label = "15,000 persons")) +
  # geom_text(aes(x = 262, y = 445, label = "7,400 persons")) +
  # geom_text(aes(x = 262, y = 430, label = "5,500 persons")) +
  labs(title = "Multiple Linear Regression",
       subtitle = "Predictor: Population",
       x = "Population",
       y = "Revenue") +
  xlim(c(50, 275))+
  theme_bw()
```
8. Put the graphs from last week and Q7 side by side (hint: use `ggpubr`).

9. Compare and contrast the modeling results and the graphs that are side by side in Q8.

10. Why do you think I assigned this particular dataset for analysis? What are you taking away from this assignment?