---
title: "Project 1"
author: "Hector Gavilanes"
format: html
self-contained: true
editor: source
---

```{r setup, echo = TRUE, warnings = FALSE, message = FALSE}
library(tidyverse)
library(gsheet)
library(formattable)
library(gridExtra)
```


**1. Members of the UWF basketball team decided to see if there was a difference in scoring on free throws during the first and second halves of games. Data was collected over thirteen games and is as follows:**

| 1st Half | 19 | 13 | 7 | 21 | 20 | 13 | 14 | 8 | 11 | 20 | 16 | 9 | 10 |
|----------|----|----|---|----|----|----|----|---|----|----|----|---|----|
| **2nd Half** | **17** | **9**  | **8** | **22** | **14** | **9** | **15** | **8** | **12** | **14** | **8**  | **8** | **10** |

```{r}
# clear environment
rm(list = ls())
# data as vector
half_1 <- c(19, 13, 7, 21, 20, 13, 14, 8, 11, 20, 16, 9, 10)
half_2 <- c(17, 9, 8, 22, 14, 9, 15, 8, 12, 14, 8, 8, 10)

# create data frame
basketball <- tibble(half_1,half_2)
# transform table from wide to long
data_long <- basketball %>% 
  drop_na() %>% 
  pivot_longer(cols = half_1:half_2, 
               names_to = "Half", 
               values_to = "Scores")
# ---CUSTOM COLORS---
custom_green_dark = "#71CA97"
custom_green_light = "#DeF7E9"
custom_gray = "#3D3D3D"
custom_orange = "#FC7307"
custom_purple = "#5A18C9"
custom_tan_light = "#F7E3D4"
custom_density_hue = c(custom_purple,custom_orange)
# ---FUNCTIONS---
# f00 - Change default table format
change_table_format <- function(data_table){
  # local variables
  first_column <- data_table[1]
  dataframe_length <- NCOL(data_table)
  # format
  formattable(data_table, align = c("l",rep("c", dataframe_length - 1)),
              list(first_column = formatter("span", style = ~ style(color = custom_gray, font.weight = "bold")),
              area(row = 1, col = 2:dataframe_length) ~ color_tile(custom_green_light, custom_green_dark)
              )#/end list
              )#/end formattable
}
# f01 - get custom summary statistics
# !! sym() creates a symbol from a string
get_custom_summary <- function(data_table, category, values){
  summary <- data_table %>%
    group_by(!! sym(category)) %>% 
    summarise(Avg = mean(!! sym(values)),
              Median = median(!! sym(values)),
              SD = sd(!! sym(values)),
              Min = min(!! sym(values)),
              Max = max(!! sym(values)),
              IQRange = IQR(!! sym(values))
              )
    return(summary)
}
```

**a. Find the appropriate summary statistics to summarize the data.**
```{r}
# summary statistics with custom table format
change_table_format(get_custom_summary(data_long, "Half","Scores"))
```

The mean for the first half is 13.92, representing 66.28% of the maximum score of 21. 
The mean for the second half is 11.85, representing 53.86% of the maximum score of 22. 
Even though the team has consistent scores, the first half is slightly better than the second. 
For both halves, the similar values for standard deviation indicate that the scores are around the mean.
We can infer that the data distribution is symmetrical because the mean values are higher than the median.

**b. Perform the appropriate *t* test to determine if there is a difference in the number of free throws made during the first and second halves of games. Remember to typeset the hypothesis test results.**
```{r}
# dependent two.sided
t.test(basketball$half_1, basketball$half_2,
       paired = TRUE)
```
### Hypothesis
$H_0$: $\mu_{1st} = \mu_{2nd}$\
$H_1$: $\mu_{1st} \ne \mu_{2nd}$\

### Test Statistics
$t_0 = 2.359$

### p-Value
$p = 0.036$

## Conclusion
Reject $H_0$ at the $\alpha$=0.05 level. There is sufficient evidence to suggest that there is a difference in the number of free throws made during the first and second halves of games.


**c. Perform the appropriate nonparametric test to determine if there is a difference in the number of free throws made during the first and second halves of games. Remember to typeset the hypothesis test results.**
```{r}
# dependent two.sided nonparametric
basket_wilcox <- wilcox.test(basketball$half_1, basketball$half_2,
            paired = TRUE,
            exact = FALSE)
```
### Hypothesis
$H_0$ : $M_{1st} = M_{2nd}$\
$H_1$ : $M_{1st} \ne M_{2nd}$\

### Test Statistics
$t_0 = 54$

### p-Value
$p = 0.065$

## Conclusion
Reject $H_0$ at the $\alpha$=0.05 level. There is sufficient evidence to suggest that there is a difference in the number of free throws made during the first and second halves of games.

**d. Construct the necessary graphs / hypothesis tests to assess the assumptions on the *t* test.**

### Assumption
```{r}
graph_paired_qq_plot <- function(data_table, after_sample, before_sample){
  data_table %>%
    mutate(diff = after_sample - before_sample) %>%
    ggplot(aes(sample = diff)) +
    stat_qq(size=3, alpha = 0.6, color = custom_orange) +
    stat_qq_line() +
    theme_bw() +
    labs(title = "Basketball Free Throws", subtitle = "QQ Plot")+
    xlab("Theoretical Free Throws") +
    ylab("Sample Free Throws")
}
graph_paired_qq_plot(basketball, half_2, half_1)
```
```{r}
# summary statistics
# median by dependent samples
summary_long_table <- data_long %>% 
  group_by(Half) %>%
  summarise_all(list(Median = median))

change_table_format(summary_long_table)

# Basketball density
basketball_density <- data_long %>% 
  group_by(Half) %>% 
  ggplot(aes(x = Scores, color = Scores, fill = Half)) +
  geom_density(data = data_long, aes(color = Half), alpha = 0.25)+
  geom_vline(data = summary_long_table, aes(xintercept=Median, color=Half), 
             linetype="dashed", size = 1.1)+
  annotate("text", x=12, y=0.025, label="Median", angle=90, size=5, color= custom_purple) +
  xlim(0,30)+
  ylab("Free Throws")+
  labs(title = "Basketball Free Throws", subtitle = "Density Plot")+
  theme_bw()+
  scale_fill_manual(values = custom_density_hue)+
  scale_color_manual(values = custom_density_hue)

```
    
**e. State the assumption(s) on the *t* test and provide your assessment based on results in part (d).**

The assumption of normality may not be valid. Normal distribution is not assumed.

**f. Based on your assessment in part (e), are you going to report the *t* test or the nonparametric test?**

The report will use the **nonparametric test**.

**g. Write a brief paragraph describing the results of your analysis. Include summary statistics, results of your chosen hypothesis test from part (f), and a graph to help convey the results.**

# UWF Basketball Free Throw Statistics Report

## Analysis
The first half has shown better scores. The team has more energy at the beginning of the match; therefore, they can make better free throws. The UWF basketball team has to improve stamina and concentration throughout the game. The team has to perform better under pressure.


### Summary Statistics

The following summary statistics displays the difference in scores.
```{r, echo=FALSE}
change_table_format(get_custom_summary(data_long, "Half","Scores"))
```
The mean for the first half is 13.92, representing 66.28% of the maximum score of 21. 
The mean for the second half is 11.85, representing 53.86% of the maximum score of 22. 
Even though the team has consistent scores, the first half is slightly better than the second. 
For both halves, the similar values for standard deviation indicate that the scores are around the mean.
We can infer that the data distribution is symmetrical because the mean values are higher than the median.

### NonParametric Test

#### Wilcoxon Test
```{r, echo=FALSE}
basket_wilcox
```
#### Hypothesis
$H_0$ : $M_{1st} = M_{2nd}$\
$H_1$ : $M_{1st} \ne M_{2nd}$\

#### Test Statistics
$t_0 = 54$

#### p-Value
$p = 0.065$

#### Conclusion
Reject $H_0$ at the $\alpha$=0.05 level. There is sufficient evidence to suggest that there is a difference in the number of free throws made during the first and second halves of games.

### Assumption
The assumption of normality may not be valid. The normal distribution is not assumed.
The following QQ and Density plots demonstrate some discrepancies between the two halves.
```{r, echo=FALSE}
graph_paired_qq_plot(basketball, half_2, half_1)
basketball_density
```



**2. It is hypothesized that students in College Algebra skip class less often than students in General Chemistry I; both courses seat 90 students. Data on the total number of absences per section was collected for one fall semester, and is as follows:**

| General Chemistry I | 161 | 140 | 136 | 171 | 106 | 149 |
|---------------------|-----|-----|-----|-----|-----|-----|
| **College Algebra** | **158** | **132** | **135** | **185** | **106** | **150** |


```{r}
# clear environment
#rm(list = ls())
# data as vector
chemistry <- c(161, 140, 136, 171, 106, 149)
algebra <- c(158, 132, 135, 185, 106, 150)

# create data frame
skip_class <- tibble(chemistry, algebra)
# transform table from wide to long
skip_long <- skip_class %>%
  drop_na() %>%
  pivot_longer(cols = chemistry: algebra,
               names_to = "Course",
               values_to = "Absence")

```


**a. Find the appropriate summary statistics to summarize the data.**
```{r}
change_table_format(get_custom_summary(skip_long, "Course", "Absence"))
```
Algebra's higher mean absences are 144.33, and Chemistry's mean absences are 143.83.
The mean for Algebra represents 78.01% of the maximum absences of 185. Chemistry's mean represents 84.11% of the maximum of 171. 
Chemistry has a lower standard deviation (26.76), indicating data points are around the mean; in contrast, Algebra has some dispersed data points because of the higher standard deviation (22.66) value. 
The Chemistry data is likely skewed in distribution because the median (144.5) is higher than the mean (143.83). On the other hand, the Algebra data is likely to be normally distributed because the median (142.5) is lower than the mean (144.33).


**b. Perform the appropriate *t* test to determine if there are fewer absences in College Algebra. Remember to typeset the hypothesis test results.**
```{r}
# two-independent sample T-test
t.test(skip_class$algebra, skip_class$chemistry,
       alternative = "less")
```

### Hypothesis
$H_0$: $\mu_{Algebra} \ge \mu_{Chemistry}$\
$H_1$: $\mu_{Algebra} < \mu_{Chemistry}$\

### Test Statistics
$t_0 = 0.035$

### p-Value
$p = 0.51$

## Conclusion
Keep $H_0$ at the $\alpha$=0.05 level. There is not sufficient evidence to determine that there are fewer absences in College Algebra.

**c. Perform the appropriate nonparametric test to determine if there are fewer absences in College Algebra. Remember to typeset the hypothesis test results.**
```{r}
wilcox.test(skip_long$Absence ~ skip_long$Course,
            alternative = "less",
            exact = FALSE)

# wilcox.test(skip_class$algebra, skip_class$chemistry,
#             alternative = "less",
#             exact = FALSE)
```

### Hypothesis
$H_0$: $M_{Algebra} \ge M_{Chemistry}$\
$H_1$: $M_{Algebra} < M_{Chemistry}$\

### Test Statistics
$t_0 = 16.5$

### p-Value
$p = 0.44$

## Conclusion
Keep $H_0$ at the $\alpha$=0.05 level. There is not sufficient evidence to determine that there are fewer absences in College Algebra.

**d. Construct the necessary graphs / hypothesis tests to assess the assumptions on the *t* test.**

### Assumptions

#### QQ Plot
```{r}
# algebra qq
algebra_qq <- skip_long %>% 
  filter(Course == "algebra") %>% 
  ggplot(aes(sample = Absence)) +
  stat_qq(size=3, alpha = 0.7, color = "magenta") +
  stat_qq_line() +
  theme_minimal() +
  xlab("Theoretical Algebra") +
  ylab("Sample Algebra")+
  ylim(90, 190)
# chemistry qq
chemistry_qq <- skip_long %>% 
  filter(Course == "chemistry") %>% 
  ggplot(aes(sample = Absence)) +
  stat_qq(size=3, alpha = 0.7, color = custom_orange) +
  stat_qq_line() +
  theme_minimal() +
  xlab("Theoretical Chemistry") +
  ylab("Sample Chemistry")+
  ylim(90, 190)
# Combine both graphs
grid.arrange(algebra_qq, chemistry_qq, nrow=1, bottom = "Courses", top = "QQ Plot of Absences by Course")
```
### Assumption Supporting Graph
```{r}
# median by courses
groups_median <- skip_long %>% 
  group_by(Course) %>% 
  summarise(Median = median(Absence))
change_table_format(groups_median)
```
```{r}
# absences density
absence_density <- skip_long %>% 
  group_by(Course) %>% 
  ggplot(aes(x = Absence, colour = Absence, fill = Course)) +
  geom_density(data = skip_long, aes(color = Course), alpha = 0.25)+
  geom_vline(data = groups_median, aes(xintercept=Median, color=Course), 
             linetype="dashed", size = 1.1)+
  annotate("text", x=138, y=0.005, label="Median", angle=90, size=6, color=custom_purple)+
  labs(title = "Absences by Course", subtitle = "Density Plot")+
  ylab("Course")+
  theme(legend.position="bottom")+
  theme_bw()+
  scale_fill_manual(values = custom_density_hue)+
  scale_color_manual(values = custom_density_hue)+
  xlim(70,220)

absence_density
```

    
**e. State the assumption(s) on the *t* test and provide your assessment based on results in part (d).**

The assumption of normality may be valid. Normal distribution is assumed. 

**f. Based on your assessment in part (e), are you going to report the *t* test or the nonparametric test?**

The report will use the **T-test**.

**g. Write a brief paragraph describing the results of your analysis. Include summary statistics, results of your chosen hypothesis test from part (f), and a graph to help convey the results.**

# Number of Absences per Course Report

## Analisys
College Algebra students skip more classes than students in General Chemistry. Algebra students need more time to understand the subject; however, students might feel frustrated. The educators should offer different tools to help students to understand the curriculum. Moreover, students can complement their studies with tutoring classes, video recordings, and study groups.  

### Summary Statistics
The following summary statistics displays the difference in absences per course.
```{r, echo=FALSE}
change_table_format(get_custom_summary(skip_long, "Course", "Absence"))
```
Algebra's higher mean absences are 144.33, and Chemistry's mean absences are 143.83.
The mean for Algebra represents 78.01% of the maximum absences of 185. Chemistry's mean represents 84.11% of the maximum of 171. 
Chemistry has a lower standard deviation (26.76), indicating data points are around the mean; in contrast, Algebra has some dispersed data points because of the higher standard deviation (22.66) value. 
The Chemistry data is likely skewed in distribution because the median (144.5) is higher than the mean (143.83). On the other hand, the Algebra data is likely to be normally distributed because the median (142.5) is lower than the mean (144.33).


### T-test
```{r, echo=FALSE}
t.test(skip_long$Absence ~ skip_long$Course,
       alternative = "less")
```

#### Hypothesis
$H_0$: $\mu_{Algebra} \ge \mu_{Chemistry}$\
$H_1$: $\mu_{Algebra} < \mu_{Chemistry}$\

#### Test Statistics
$t_0 = 0.035$

#### p-Value
$p = 0.51$

### Conclusion
Keep $H_0$ at the $\alpha$=0.05 level. There is not sufficient evidence to determine that there are fewer absences in College Algebra.

### Assumption
The assumption of normality may be valid. The normal distribution is assumed. The following QQ and Density plots demonstrate how the data match the general appearance of a normal distribution.

```{r, echo=FALSE}
grid.arrange(algebra_qq, chemistry_qq, nrow=1, bottom = "Courses", top = "QQ Plot of Absences by Course")
absence_density
```

